### 2015-10-26 report/paper.tex
Add report files

\documentclass{article} % For LaTeX2e
% We will use NIPS submission format
\usepackage{nips13submit_e,times}
% for hyperlinks
\usepackage{hyperref}
\usepackage{url}
% For figures
\usepackage{graphicx} 
\usepackage{subfigure} 
% math packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsopn}
\usepackage{ifthen}
\usepackage{natbib}

\title{Project-I by Group Sao Paulo}

\author{
Damien Engels\\
EPFL \\
\texttt{damien.engels@epfl.ch} \And No\'emie Jaquier\\
EPFL \\
\texttt{noemie.jaquier@epfl.ch} \\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\nipsfinalcopy

\begin{document}

\maketitle

\begin{abstract}
In this report, ...
\end{abstract}

\section{Data Description}
Our train-data consists of ...

\section{Data visualization and cleaning}
We performed basic exploratory data analysis on our data. Figure \ref{fig:boxplotX} ...

\begin{figure}[!h] % !t
\center
\subfigure[Boxplot of real-valued $\mathbf{X}$. Data is not centered and therefore we normalize it.]{\includegraphics[width=2.5in]{figures/boxplotX.pdf} \label{fig:boxplotX}}
\hfill
\subfigure[Histogram of $\mathbf{y}$. We can clearly see two outliers.]{\includegraphics[width=2.5in]{figures/histY.pdf} \label{fig:histY}}
\caption{}
\end{figure}

\section{Ridge regression}
We applied least-squares and ridge regression to this dataset...

\section{Feature transformations}
We tried several feature transformations...

\section{Summary}
In this report, we ...

\subsubsection*{Acknowledgments}
...

\subsubsection*{References}

\end{document}

### 2015-10-28 methods/penLogisticRegression.m
Experiments with regression

@@ -32,7 +32,7 @@ for k = 1:maxIters
     % Look at the convergence
     if k>1
         fprintf('%d : %f\n', k, max(abs(beta-beta_all(:,k-1))));
-        if abs(beta-beta_all(:,k-1))<0.01
+        if abs(beta-beta_all(:,k-1))<0.001
             % If the difference between last and present beta is small...
             % break
             break;

### 2015-10-28 methods/computeRMSE.m
Modularised the code and modify clustering

@@ -1,0 +1,8 @@
+function [ error ] = computeRMSE( y, yHat )
+%computeRMSE
+
+N = length(y);
+error = sqrt(sum((y - yHat).^2)./N);
+
+end
+

### 2015-10-28 methods/crossValidationRegression.m

@@ -1,0 +1,21 @@
+function [ trRMSE, teRMSE ] = crossValidationRegression( y, X, groups )
+%crossValidationRegression
+
+trRMSE = [];
+teRMSE = [];
+
+for groupNumber = 1:groups
+    %% Separe X and y in train and test parts
+    [ XTr, XTe, yTr, yTe ] = divideDataSet( X, y, groups, groupNumber );
+    
+    %% Make prediction for test data
+    yTrPred = predict( XTr, yTr, XTr );
+    yTePred = predict( XTr, yTr, XTe );
+    
+    trRMSE = [trRMSE computeRMSE(yTr, yTrPred)];
+    teRMSE = [teRMSE computeRMSE(yTe, yTePred)];
+    
+end
+
+end
+

### 2015-10-28 methods/leastSquaresGD.m

@@ -1,7 +1,7 @@
 function beta = leastSquaresGD(y,tX,alpha)
 
 % Maximum number of iterations
-maxIters = 1000; % TO DEFINE
+maxIters = 5000; % TO DEFINE
 epsilon = 0.01;
 
 % Number of data

### 2015-10-28 methods/mainRegression.m

@@ -1,0 +1,7 @@
+close all
+clear
+clc
+
+load('../data/SaoPaulo_regression.mat')
+
+[ trRMSE, teRMSE ] = crossValidationRegression( y_train, X_train, 4 );

### 2015-10-28 methods/predict.m

@@ -1,0 +1,75 @@
+function [ yPred ] = predict( XTr, yTr, XTe )
+%predict
+
+%% Clustering of the data with kMeans
+% Need to do that with non-normalised data !
+X = XTr;
+y = yTr;
+[cluster, centers] = kmeans([X(:,55) y], 2);
+[subcluster, subcenters] = kmeans([X(cluster==1,6) y(cluster==1)], 2);
+
+finalCluster = cluster;
+finalCluster(cluster == 2) = 3;
+finalCluster(cluster==1) = subcluster;
+finalCenters = [subcenters(1,:);centers(2,:);subcenters(2,:)];
+
+%% Normalise the data (train)
+
+[XTrNormalised, meanX, stdX] = normalise(XTr);
+[yTrNormalised, meanY, stdY] = normalise(yTr);
+
+%% Remove outliers (train)
+
+%[XFiltered, yFiltered] = removeOutlierLines(XNormalised, yNormalised, 3, 2);
+XTrFiltered = fixOutliers(XTrNormalised, 3);
+yTrFiltered = yTrNormalised;
+
+%% Remove correlated columns using PCA (test)
+[XTrKept, V, VReduced] = pca(XTrFiltered, 1);
+
+%% Preprocess test data
+% Normalise given mean and std of the training set
+XTeNormalised = (XTe - ones(size(XTe,1),1)*meanX)./(ones(size(XTe,1),1)*stdX);
+% Reduce the dimension given PCA of training set
+XTeNormalised = (VReduced'*(V*XTeNormalised'))';
+
+%% Compute tX for train and test
+tXTr = [ones(length(XTrKept), 1)  XTrKept];
+tXTe = [ones(length(XTeNormalised), 1)  XTeNormalised];
+
+%% Train for clusters using minimum distance to the center in 55th dimension
+clusterPred = findClosestGroup([XTe(:,55)], centers(:,1));
+finalClusterPred = clusterPred;
+finalClusterPred(clusterPred == 2) = 3;
+
+subClusterPred = findClosestGroup([XTe(finalClusterPred~=3,6)], finalCenters(1:2,1));
+finalClusterPred(finalClusterPred~=3) = subClusterPred;
+
+figure()
+plot(XTe(finalClusterPred==1,55),XTe(finalClusterPred==1,6),'*');
+hold on
[…]

### 2015-10-28 methods/preprocessing.m

@@ -20,7 +20,7 @@ plot(X_train(:,55),y_train,'*');
 binaryDim = [50];
 threeDim = [20 21 26 37 46];
 fourDim = [32 45 51 65];
-twoCloudsDim = [55]; 
+twoCloudsDim = [6 55]; 
 
 %% Cross-validation
 datagroups = 3;
@@ -35,7 +35,7 @@ group = 2;
     X = XTr;
     y = yTr;
     [cluster, centers] = kmeans([X(:,55) y], 2);
-    [subcluster, subcenters] = kmeans([X(cluster==1,55) y(cluster==1)], 2);
+    [subcluster, subcenters] = kmeans([X(cluster==1,6) y(cluster==1)], 2);
     
     finalcluster = cluster;
     finalcluster(cluster == 2) = 3;
@@ -59,6 +59,15 @@ group = 2;
     plot(X(finalcluster==3,55),y(finalcluster==3,1),'*');
     plot(XTe(:,55),yTe(:,1),'*');
     plot(finalcenters(:,1), finalcenters(:,end),'.r', 'markersize',30);
+    legend('cluster 1','cluster 2','cluster 3','test data','finalcenters');
+    
+    figure()
+    plot3(X(finalcluster==1,6),X(finalcluster==1,55),y(finalcluster==1,1),'*');
+    hold on
+    plot3(X(finalcluster==2,6),X(finalcluster==2,55),y(finalcluster==2,1),'*');
+    plot3(X(finalcluster==3,6),X(finalcluster==3,55),y(finalcluster==3,1),'*');
+    plot3(XTe(:,6),XTe(:,55),yTe(:,1),'*');
+%     plot3(finalcenters(:,1),centers(:,2), finalcenters(:,end),'.r', 'markersize',30);
     legend('cluster 1','cluster 2','cluster 3','test data','finalcenters');
     %% Normalise the data (train)
     
@@ -104,10 +113,11 @@ group = 2;
     
     beta = leastSquaresGD(y(finalcluster~=3), tXTr(finalcluster~=3,:), 0.01);
     yPred = (tXTe(finalclusterPred~=3,:) * beta)*stdY + meanY;
-    betaCl = logisticRegression( cluster(cluster~=3), [XTr(cluster~=3,55) yTr(cluster~=3)], 0.001 );
-    subclusterPred =  sigma([XTe(finalclusterPred~=3,55) yPred]* betaCl);
[…]

### 2015-10-28 methods/kmeans.m
Change kmeans method

@@ -1,50 +1,0 @@
-function [ groups, centers ] = kmeans( X, k )
-%KMEANS Clustering X data in k clusters
-
-dim = size(X, 2);
-len = length(X);
-
-% Pre sort
-norms = zeros(len, 1);
-for line = 1:len
-    norms(line) = norm(X(line));
-end
-[~, I] = sort(norms);
-
-presorted = X(I, :);
-
-
-% Inititialize centers
-centers = zeros(k, dim);
-for cluster = 1:k
-   sizeChunk = floor(len/k);
-   start = (cluster - 1) * sizeChunk + 1;
-   ennd = min(cluster * sizeChunk + 1, len);
-   centers(cluster, :) = mean(presorted(start:ennd ,:));
-end
-
-groups = zeros(len, 1);
-old_centers = zeros(k, dim);
-while centers ~= old_centers
-
-    % Recompute cluster membership
-    for line = 1:len
-       distances = zeros(k, 1);
-       for cluster = 1:k
-           distances(cluster) = norm(centers(cluster, :) - X(line, :));
-       end
-       group = find(distances == min(distances));
-       groups(line) = group;
-    end
-   
-    
-    % Recompute cluster centers
-    old_centers = centers;
-    for cluster = 1:k
-        centers(cluster, :) = mean(X(groups == cluster, :));
-    end  
-    
-end
-
-end
-

### 2015-10-28 methods/kmeansStable.m

@@ -1,0 +1,31 @@
+function [ groups, centers ] = kmeansStable( X, k )
+%KMEANS Clustering X data in k clusters
+% We need this function because the native MATLAB kmeans algorithm uses 
+% random heuristics to initialise the centers
+
+dim = size(X, 2);
+len = length(X);
+
+% Pre sort
+norms = zeros(len, 1);
+for line = 1:len
+    norms(line) = norm(X(line));
+end
+[~, I] = sort(norms);
+
+presorted = X(I, :);
+
+
+% Inititialize centers
+centers = zeros(k, dim);
+for cluster = 1:k
+   sizeChunk = floor(len/k);
+   start = (cluster - 1) * sizeChunk + 1;
+   ennd = min(cluster * sizeChunk + 1, len);
+   centers(cluster, :) = mean(presorted(start:ennd, :));
+end
+
+[groups, centers] = kmeans(X, k, 'Start', centers);
+
+end
+

### 2015-10-28 methods/predict.m

@@ -1,12 +1,12 @@
 function [ yPred ] = predict( XTr, yTr, XTe )
 %predict
 
-%% Clustering of the data with kMeans
+%% Clustering of the data with kmeansStable
 % Need to do that with non-normalised data !
 X = XTr;
 y = yTr;
-[cluster, centers] = kmeans([X(:,55) y], 2);
-[subcluster, subcenters] = kmeans([X(cluster==1,6) y(cluster==1)], 2);
+[cluster, centers] = kmeansStable([X(:,55) y], 2);
+[subcluster, subcenters] = kmeansStable([X(cluster==1,6) y(cluster==1)], 2);
 
 finalCluster = cluster;
 finalCluster(cluster == 2) = 3;

### 2015-10-28 methods/kmeansStable.m
Add cool statistic stuff

@@ -1,4 +1,4 @@
-function [ groups, centers ] = kmeansStable( X, k )
+function [ groups, centers, sumd ] = kmeansStable( X, k )
 %KMEANS Clustering X data in k clusters
 % We need this function because the native MATLAB kmeans algorithm uses 
 % random heuristics to initialise the centers
@@ -25,7 +25,7 @@ for cluster = 1:k
    centers(cluster, :) = mean(presorted(start:ennd, :));
 end
 
-[groups, centers] = kmeans(X, k, 'Start', centers);
+[groups, centers, sumd] = kmeans(X, k, 'Start', centers);
 
 end
 

### 2015-10-28 methods/cardinalities.m
Add cool statistic stuff

@@ -1,0 +1,14 @@
+function [ res ] = cardinalities( X )
+%CARDINALITIES Counts the cardinality of each columns
+%   Returns for each column, the count of unique elements it has
+
+dims = size(X, 2);
+
+res = zeros(1, dims);
+for i =  1:dims
+    res(i) = length(unique(X(:, i)));
+end
+
+
+end
+

### 2015-10-28 methods/plotData.m

@@ -1,0 +1,13 @@
+%% Plot all the dimensions of X_train to visually inspect
+clear
+clc
+
+load('../data/SaoPaulo_regression.mat');
+len = length(X_train);
+dim = size(X_train, 2);
+
+for d = 1:dim
+   figure();
+   hold on;
+   plot(X_train(:, d), y_train, '*');
+end

### 2015-10-28 methods/statistics.m

@@ -1,0 +1,46 @@
+%% Clean workspace & Load data
+clear
+clc
+
+load('../data/SaoPaulo_regression.mat');
+len = length(X_train);
+dim = size(X_train, 2);
+
+section = @(text) fprintf('\n\n%s\n%s\n', text, repmat('=', 1, length(text)));
+
+
+%% Cardinalities
+section('Cardinalities');
+
+cards = cardinalities(X_train);
+
+for card = unique(cards)
+    dims = find(cards == card);
+    fprintf('%2d dimensions with cardinality %d: ', length(dims), card);
+    if(card < len)
+        disp(dims);
+    else
+        fprintf('All other dimesions\n');
+    end
+end
+
+%% Clusters
+section('2-way Clusters');
+
+sums = zeros(1, dim);
+for d = 1:dim
+   column = X_train(:, d);
+   [idx, C, sumd] = kmeansStable(column, 2);
+   sums(d) = norm(sumd);
+end
+
+% Normalize
+sums = normalise(sums);
+
+% Find outliers ????
+dimsClustered = find(sums > 2);
+
+fprintf('Dimensions with well defined 2-way clusters:');
+disp(dimsClustered);
+
+

### 2015-10-28 methods/mainRegression.m
Make adaptations to new kmeans

@@ -4,4 +4,4 @@ clc
 
 load('../data/SaoPaulo_regression.mat')
 
-[ trRMSE, teRMSE ] = crossValidationRegression( y_train, X_train, 4 );
+[ trRMSE, teRMSE ] = crossValidationRegression( y_train, X_train, 3 );

### 2015-10-28 methods/predict.m

@@ -5,8 +5,8 @@ function [ yPred ] = predict( XTr, yTr, XTe )
 % Need to do that with non-normalised data !
 X = XTr;
 y = yTr;
-[cluster, centers] = kmeansStable([X(:,55) y], 2);
-[subcluster, subcenters] = kmeansStable([X(cluster==1,6) y(cluster==1)], 2);
+[cluster, centers, ~] = kmeansStable([X(:,55) y], 2);
+[subcluster, subcenters, ~] = kmeansStable([X(cluster==1,6) y(cluster==1)], 2);
 
 finalCluster = cluster;
 finalCluster(cluster == 2) = 3;

### 2015-10-30 methods/correlation.m
Refactoring

@@ -1,6 +1,6 @@
 function [ I, corrs ] = correlation( X, y )
 %CORRELATION Returns the indices with highest correlation between input and output 
-%   corrs returns the correlation
+%   corrs returns the correlation value for each dimension
 
 dim = size(X, 2);
 

### 2015-10-30 methods/crossValidation.m

@@ -1,0 +1,25 @@
+function [ trRMSE, teRMSE ] = crossValidation( X, y, groups, predict)
+%crossValidationRegression Returns RMSE for train and test for this prediciton function
+% X         Data input
+% y         Output
+% groups    Number of groups to form
+% predict   The prediciton function
+
+trRMSE = zeros(1, groups);
+teRMSE = zeros(1, groups);
+
+for groupNumber = 1:groups
+    %% Separe X and y in train and test parts
+    [ XTr, XTe, yTr, yTe ] = divideDataSet( X, y, groups, groupNumber );
+    
+    %% Make prediction for test data
+    yTrPred = predict( XTr, yTr, XTr );
+    yTePred = predict( XTr, yTr, XTe );
+    
+    trRMSE(groupNumber) = computeRMSE(yTr, yTrPred);
+    teRMSE(groupNumber) = computeRMSE(yTe, yTePred);
+    
+end
+
+end
+

### 2015-10-30 methods/crossValidationRegression.m

@@ -1,21 +1,0 @@
-function [ trRMSE, teRMSE ] = crossValidationRegression( y, X, groups )
-%crossValidationRegression
-
-trRMSE = [];
-teRMSE = [];
-
-for groupNumber = 1:groups
-    %% Separe X and y in train and test parts
-    [ XTr, XTe, yTr, yTe ] = divideDataSet( X, y, groups, groupNumber );
-    
-    %% Make prediction for test data
-    yTrPred = predict( XTr, yTr, XTr );
-    yTePred = predict( XTr, yTr, XTe );
-    
-    trRMSE = [trRMSE computeRMSE(yTr, yTrPred)];
-    teRMSE = [teRMSE computeRMSE(yTe, yTePred)];
-    
-end
-
-end
-

### 2015-10-30 methods/crossValidationParam.m
Refactored functional style

@@ -1,0 +1,24 @@
+function [ trRMSE, teRMSE, param ] = crossValidationParam( X, y, groups, predictParam)
+%crossValidationRegression
+
+param = logspace(-5, 8, 20);
+lenParam = length(param);
+
+trRMSE = zeros(1, lenParam);
+teRMSE = zeros(1, lenParam);
+
+for paramIdx = 1:lenParam
+    
+    lambda = param(paramIdx);
+    
+    predict = @(Xtr, ytr, Xte) predictParam(Xtr, ytr, Xte, lambda);
+    
+    [trError, teError] = crossValidation(X, y, groups, predict);
+    
+    trRMSE(paramIdx) = mean(trError);
+    teRMSE(paramIdx) = mean(teError);
+    
+end
+
+end
+

### 2015-10-30 methods/crossValidationParamRegression.m

@@ -1,30 +1,0 @@
-function [ trRMSE, teRMSE, param ] = crossValidationParamRegression( y, X, groups )
-%crossValidationRegression
-
-trRMSE = [];
-teRMSE = [];
-
-param = logspace(-5,8,20);
-
-for paramNumber = 1:length(param)
-    trError = [];
-    teError = [];
-    lambda = param(paramNumber);
-    for groupNumber = 1:groups
-        %% Separe X and y in train and test parts
-        [ XTr, XTe, yTr, yTe ] = divideDataSet( X, y, groups, groupNumber );
-        
-        %% Make prediction for test data
-        yTrPred = predict( XTr, yTr, XTr, lambda );
-        yTePred = predict( XTr, yTr, XTe, lambda );
-        
-        trError = [trError computeRMSE(yTr, yTrPred)];
-        teError = [teError computeRMSE(yTe, yTePred)];
-    end
-    trRMSE = [trRMSE mean(trError)];
-    teRMSE = [teRMSE mean(teError)];
-    
-end
-
-end
-

### 2015-10-30 methods/divideDataSet.m

@@ -4,7 +4,7 @@ function [ XTr, XTe, yTr, yTe ] = divideDataSet( X, y, K, k )
 N = size(y,1);
 idx = randperm(N);
 Nk = floor(N/K);
-%idxCV = zeros(K, N);
+idxCV = zeros(K, Nk);
 
 for k = 1:K
     idxCV(k, :) = idx(1+(k-1)*Nk:k*Nk);

### 2015-10-30 methods/mainRegression.m

@@ -4,7 +4,7 @@ clc
 
 load('../data/SaoPaulo_regression.mat')
 
-% [ trRMSE, teRMSE ] = crossValidationRegression( y_train, X_train, 3 );
-predict = @(Xtr, Ytr, Xte) predictRegression(Xtr, Ytr, Xte, 10^5);
+% predict = @(Xtr, Ytr, Xte) predictRegression(Xtr, Ytr, Xte, 10^5);
+% [ trRMSE, teRMSE ] = crossValidation(y_train, X_train, 3, predict);
 
-[ trRMSE, teRMSE ] = crossValidation( X_train, y_train, 3, predict);
+[ trRMSE, teRMSE, lambda ] = crossValidationParam(X_train, y_train, 3, @predictRegression);

### 2015-10-30 methods/crossValidationParamRegression.m
Update regression and classification

@@ -4,7 +4,7 @@ function [ trRMSE, teRMSE, param ] = crossValidationParamRegression( y, X, group
 trRMSE = [];
 teRMSE = [];
 
-param = logspace(-5,8,20);
+param = logspace(-5,-2,10);
 
 for paramNumber = 1:length(param)
     trError = [];

### 2015-10-30 methods/penLogisticRegression.m

@@ -8,14 +8,10 @@ N = size(tX,1);
 
 % Beta initialisation
 beta = zeros(size(tX,2),1);
-beta = zeros(size(tX,2),1);
 
 % Iterate the gradient descent
 for k = 1:maxIters
     
-    % Compute cost of logistic regression
-    L = computeCostPenLogisticReg(y, tX, beta, lambda);
-    fprintf('%d : %f\n', k,L);
     % Compute the gradient
     g = computePenGradient(y, tX, beta, lambda);
     % Compute the Hessian
@@ -25,6 +21,11 @@ for k = 1:maxIters
     % beta = beta - alpha.*g;
     % Update beta using Newton's method
     beta = beta - alpha.*(H^-1)*g;
+    
+    % Compute cost of logistic regression
+    L = computeCostPenLogisticReg(y, tX, beta, lambda);
+    alpha = abs(L*10^(-4));
+    fprintf('%d : %f  %f\n', k,L, alpha);
     
     % Store beta and L
     beta_all(:,k) = beta;
@@ -36,9 +37,6 @@ for k = 1:maxIters
         if abs(beta-beta_all(:,k-1))<0.001
             % If the difference between last and present beta is small...
             % break
-            break;
-        end
-        if L <0
             break;
         end
     end

### 2015-10-30 methods/predictSave2910.m
deleted stupid files

@@ -1,85 +1,0 @@
-function [ yPred ] = predict( XTr, yTr, XTe )
-%predict
-
-%% Find cardinalities of data
-cards = cardinalities(XTr);
-clusteredDim = [55, 6];
-
-%% Remove dimensions with discrete arguments
-XTr = XTr(:,cards == size(XTr,1));
-XTe = XTe(:,cards == size(XTr,1));
-clusteredDim = [46, 6];
-
-%% Clustering of the data with kmeansStable
-% Need to do that with non-normalised data !
-X = XTr;
-y = yTr;
-[cluster, centers, ~] = kmeansStable([X(:,clusteredDim(1)) y], 2);
-[subcluster, subcenters, ~] = kmeansStable([X(cluster==1,clusteredDim(2)) y(cluster==1)], 2);
-
-finalCluster = cluster;
-finalCluster(cluster == 2) = 3;
-finalCluster(cluster==1) = subcluster;
-finalCenters = [subcenters(1,:);centers(2,:);subcenters(2,:)];
-
-%% Normalise the data (train)
-
-[XTrNormalised, meanX, stdX] = normalise(XTr);
-[yTrNormalised, meanY, stdY] = normalise(yTr);
-
-%% Remove outliers (train)
-
-XTrFiltered = fixOutliers(XTrNormalised, 3);
-yTrFiltered = yTrNormalised;
-% [XTrFiltered, yFiltered] = removeOutlierLines(XTrNormalised, [yTrNormalised finalCluster], 3, 1);
-% finalCluster = yFiltered(:,2);
-% yTrFiltered = yFiltered(:,1);
-%% Remove correlated columns using PCA (test)
-[XTrKept, V, VReduced] = pca(XTrFiltered, 1);
-
-%% Preprocess test data
-% Normalise given mean and std of the training set
-XTeNormalised = (XTe - ones(size(XTe,1),1)*meanX)./(ones(size(XTe,1),1)*stdX);
-% Reduce the dimension given PCA of training set
-XTeNormalised = (VReduced'*(V*XTeNormalised'))';
-
-%% Compute tX for train and test
-tXTr = [ones(length(XTrKept), 1)  XTrKept];
-tXTe = [ones(length(XTeNormalised), 1)  XTeNormalised];
-
-%% Train for clusters using minimum distance to the center in 55th dimension
[…]

### 2015-10-30 methods/classification/plotDataClassification.m
Update regression and classification files

@@ -9,7 +9,7 @@ load('../data/SaoPaulo_classification.mat');
 keep = zeros(1, length(X_train));
 for i = 1:length(X_train) 
     line = X_train(i, :);
-    if(line(20) == 2)
+    if(line(20) == 1)
         keep(i) = 0;
     else 
         keep(i) = 1;
@@ -25,9 +25,9 @@ dim = size(X_train, 2);
 
 
 for d = 1:dim
-   %figure();
-   %hold on;
-   %plot(X_train(y_train == -1, 30), X_train(y_train == -1, d), 'r*');
-   %plot(X_train(y_train == 1, 30), X_train(y_train == 1, d), 'b*');
+   figure();
+   hold on;
+   plot(X_train(y_train == -1, 30), X_train(y_train == -1, d), 'r*');
+   plot(X_train(y_train == 1, 30), X_train(y_train == 1, d), 'b*');
    %histogram(X_train(:, d))
 end

### 2015-10-30 methods/predictRegression.m

@@ -79,8 +79,7 @@ end
 yPred = zeros(size(XTe,1),1);
 for cl = 1:size(finalCenters,1)
     % Take the data from one cluster
-    X = XTr(finalClusterPred == cl,:);
-    y = yTr(finalClusterPred == cl,:);
+    X = XTe(finalClusterPred == cl,:);
     % Normalise given mean and std of the training set
     XTeNormalised = (X - ones(size(X,1),1)*meanX(cl,:))./(ones(size(X,1),1)*stdX(cl,:));
     

### 2015-10-30 methods/preprocessingClassification.m
Update classification predictions

@@ -6,7 +6,7 @@ load('../data/SaoPaulo_classification.mat')
 
 
 %% Separe X_train in train and test to make cross-validation
-[ XTr, XTe, yTr, yTe ] = divideDataSet( X_train, yTrFiltered, 4, 2 );
+[ XTr, XTe, yTr, yTe ] = divideDataSet( X_train, y_train, 4, 2 );
 
 % cards = cardinalities( XTr );
 % XTr = XTr(:,cards < size(XTr,1));
@@ -17,13 +17,14 @@ cluster = XTr(:,20);
 clusterPred = XTe(:,20);
 
 %% Train regression model for each cluster
-V = cell(size(finalCenters,1),1);
-VReduced = cell(size(finalCenters,1),1);
-tXTr = cell(size(finalCenters,1),1);
-yTrF = cell(size(finalCenters,1),1);
-beta = cell(size(finalCenters,1),1);
+sizeCluster = length(unique(cluster));
+V = cell(sizeCluster,1);
+VReduced = cell(sizeCluster,1);
+tXTr = cell(sizeCluster,1);
+yTrF = cell(sizeCluster,1);
+beta = cell(sizeCluster,1);
 
-for cl = 1:length(unique(cluster))
+for cl = 1:sizeCluster
     % Take the data for one cluster
     X = XTr(cluster == cl,:);
     y = yTr(cluster == cl,:);
@@ -38,7 +39,7 @@ for cl = 1:length(unique(cluster))
 %     XFiltered = fixOutliers(XTrNormalised, 3);
 
     % Remove correlated columns using PCA 
-    [XTrKept, V{cl,1}, VReduced{cl,1}] = pca(XTrFiltered, 1);
+%     [XTrKept, V{cl,1}, VReduced{cl,1}] = pca(XTrFiltered, 1);
     XTrKept = XTrFiltered;
     
     % Compute tX and final y
@@ -50,32 +51,29 @@ for cl = 1:length(unique(cluster))
 
 end
 
-len = length(XFiltered);
-width = size(XFiltered, 2);
-
 %% Test
 yPred = zeros(size(XTe,1),1);
 for cl = 1:length(unique(clusterPred))
+    % Take the data from one cluster
+    X = XTe(finalClusterPred == cl,:);
     
-% Normalise given mean and std of the training set
[…]

### 2015-10-31 methods/classification/predictClassification.m
Refactoring

@@ -1,5 +1,6 @@
-function [ groups ] = predictClass( tX, beta, k )
+function [ groups ] = predictClassification( tX, beta, k )
 %predictClass Return the most probable class for each data
+
 
 for class = 1:k
     

### 2015-10-31 methods/crossValidationParam.m

@@ -1,7 +1,7 @@
-function [ trRMSE, teRMSE, param ] = crossValidationParam( X, y, groups, predictParam)
+function [ trRMSE, teRMSE, param ] = crossValidationParam( X, y, groups, predictParam, nbIter)
 %crossValidationRegression
 
-param = logspace(-5, 8, 20);
+param = logspace(-6, 10, 60);
 lenParam = length(param);
 
 trRMSE = zeros(1, lenParam);
@@ -10,13 +10,24 @@ teRMSE = zeros(1, lenParam);
 for paramIdx = 1:lenParam
     
     lambda = param(paramIdx);
-    
     predict = @(Xtr, ytr, Xte) predictParam(Xtr, ytr, Xte, lambda);
     
-    [trError, teError] = crossValidation(X, y, groups, predict);
+    trMeanError = zeros(1, nbIter);
+    teMeanError = zeros(1, nbIter);
     
-    trRMSE(paramIdx) = mean(trError);
-    teRMSE(paramIdx) = mean(teError);
+    for iter = 1:nbIter
+        [trError, teError] = crossValidation(X, y, groups, predict);
+        trMeanError(iter) = mean(trError);
+        teMeanError(iter) = mean(teError);
+    end
+    
+    trMeanError = trMeanError( abs(normalise(trMeanError)) < 2 );
+    teMeanError = teMeanError( abs(normalise(teMeanError)) < 2 );
+
+    trRMSE(paramIdx) = mean(trMeanError);
+    teRMSE(paramIdx) = mean(teMeanError);
+    
+    fprintf('Finished %d out of %d\n', paramIdx, lenParam);
     
 end
 

### 2015-10-31 methods/findClosestGroup.m

@@ -1,6 +1,7 @@
 function [ groups ] = findClosestGroup( X, centers )
 %findClosestGroup Find group of the data by computing the minimum distance
 % to the centers of the groups
+
 for line = 1:length(X)
     distances = zeros(size(centers,1), 1);
     for group = 1:size(centers,1)

### 2015-10-31 methods/fixOutliers.m

@@ -1,5 +1,8 @@
 function [ XFixed ] = fixOutliers( X, limitDeviation )
 %fixOutliers Gets the outliers back to a few standard deviations away
+%
+% limitDeviation    Number of std deviations away that makes a point an
+%                   outlier
 
 stdX = std(X);
 meanX = mean(X);

### 2015-10-31 methods/kmeansStable.m

@@ -1,5 +1,6 @@
 function [ groups, centers, sumd ] = kmeansStable( X, k )
 %KMEANS Clustering X data in k clusters
+%
 % We need this function because the native MATLAB kmeans algorithm uses 
 % random heuristics to initialise the centers
 
@@ -14,7 +15,6 @@ end
 [~, I] = sort(norms);
 
 presorted = X(I, :);
-
 
 % Inititialize centers
 centers = zeros(k, dim);

### 2015-10-31 methods/leastSquaresGD.m

@@ -1,4 +1,4 @@
-function beta = leastSquaresGD(y,tX,alpha)
+function beta = leastSquaresGD( y, tX, alpha )
 
 % Maximum number of iterations
 maxIters = 10000; % TO DEFINE

### 2015-10-31 methods/logisticRegression.m

@@ -1,4 +1,4 @@
-function beta = logisticRegression(y,tX,alpha)
+function beta = logisticRegression( y, tX, alpha )
 
 % Maximum number of iterations
 maxIters = 1000; % TO DEFINE

### 2015-10-31 methods/preprocessing.m

@@ -1,180 +1,0 @@
-close all
-clear
-clc
-
-load('../data/SaoPaulo_regression.mat')
-
-%% Visualise the data
-figure()
-histogram(y_train);
-
-figure()
-plot(X_train(:,55),y_train,'*');
-% About X data:
-% Binary dimensions :50
-% Three values dimensions : 20, 21, 26, 37, 46
-% Four values dimensions : 32, 45, 51, 65
-% Two clouds : 55
-
-% refactor to use find
-binaryDim = [50];
-threeDim = [20 21 26 37 46];
-fourDim = [32 45 51 65];
-twoCloudsDim = [6 55]; 
-
-%% Cross-validation
-datagroups = 3;
-group = 2;
-% for group = 1:datagroups
-    
-    %% Separe X_train in train and test to make cross-validation
-    [ XTr, XTe, yTr, yTe ] = divideDataSet( X_train, y_train, datagroups, group );
-    
-    %% Clustering of the data with kMeans
-    % Need to do that with non-normalised data !
-    X = XTr;
-    y = yTr;
-    [cluster, centers] = kmeans([X(:,55) y], 2);
-    [subcluster, subcenters] = kmeans([X(cluster==1,6) y(cluster==1)], 2);
-    
-    finalcluster = cluster;
-    finalcluster(cluster == 2) = 3;
-    finalcluster(cluster==1) = subcluster;
-    finalcenters = [subcenters(1,:);centers(2,:);subcenters(2,:)];
-    
-    figure()
-    plot(X(cluster==1,55),y(cluster==1,1),'*');
-    hold on
-    plot(X(cluster==2,55),y(cluster==2,1),'*');
-    plot(XTe(:,55),yTe(:,1),'*');
-    plot(centers(:,1), centers(:,end),'.', 'markersize',30);
-    plot(subcenters(:,1), subcenters(:,end),'.', 'markersize',30);
-    plot(finalcenters(:,1), finalcenters(:,end),'.r', 'markersize',30);
-    legend('cluster 1','cluster 2','test data','centers','subcenters','finalcenters');
-    
-    figure()
-    plot(X(finalcluster==1,55),y(finalcluster==1,1),'*');
-    hold on
-    plot(X(finalcluster==2,55),y(finalcluster==2,1),'*');
-    plot(X(finalcluster==3,55),y(finalcluster==3,1),'*');
-    plot(XTe(:,55),yTe(:,1),'*');
[…]

### 2015-10-31 methods/regression/mainRegression.m

@@ -7,4 +7,8 @@ load('../data/SaoPaulo_regression.mat')
 % predict = @(Xtr, Ytr, Xte) predictRegression(Xtr, Ytr, Xte, 10^5);
 % [ trRMSE, teRMSE ] = crossValidation(y_train, X_train, 3, predict);
 
-[ trRMSE, teRMSE, lambda ] = crossValidationParam(X_train, y_train, 3, @predictRegression);
+[ trRMSE, teRMSE, lambda ] = crossValidationParam(X_train, y_train, 3, @predictRegression, 4);
+
+semilogx(lambda, teRMSE, 'r*-')
+hold on
+semilogx(lambda, trRMSE, 'b*-')

### 2015-10-31 methods/regression/predictRegression.m

@@ -79,7 +79,8 @@ end
 yPred = zeros(size(XTe,1),1);
 for cl = 1:size(finalCenters,1)
     % Take the data from one cluster
-    X = XTe(finalClusterPred == cl,:);
+    X = XTr(finalClusterPred == cl,:);
+    y = yTr(finalClusterPred == cl,:);
     % Normalise given mean and std of the training set
     XTeNormalised = (X - ones(size(X,1),1)*meanX(cl,:))./(ones(size(X,1),1)*stdX(cl,:));
     

### 2015-10-31 methods/removeOutlierLines.m

@@ -1,5 +1,11 @@
 function [XFiltered, yFiltered] = removeOutlierLines(X, y, limitDeviation, limitPerLine)
-%removeOutlierLines Removes lines that contain too many outliers
+% removeOutlierLines Removes lines that contain too many outliers
+%
+% limitDeviation    The number of standard deviations away a point is to be
+%                   considered an outlier
+%
+% limitPerLine      The number of outliers per line that cause us to remove
+%                   the line
 
 stdX = std(X);
 meanX = mean(X);

### 2015-10-31 methods/regression/predictRegression.m
Fixed predictRegression

@@ -79,8 +79,8 @@ end
 yPred = zeros(size(XTe,1),1);
 for cl = 1:size(finalCenters,1)
     % Take the data from one cluster
-    X = XTr(finalClusterPred == cl,:);
-    y = yTr(finalClusterPred == cl,:);
+    X = XTe(finalClusterPred == cl,:);
+    %y = yTr(finalClusterPred == cl,:);
     % Normalise given mean and std of the training set
     XTeNormalised = (X - ones(size(X,1),1)*meanX(cl,:))./(ones(size(X,1),1)*stdX(cl,:));
     

### 2015-10-31 src/regression/predictRegression.m
Factored Cluster predicition out of main prediction file

@@ -1,5 +1,5 @@
 function [ yPred ] = predictRegression( XTr, yTr, XTe, param )
-%predict
+%predictRegression Prediction function for the regression
 
 %% Find cardinalities of data
 cards = cardinalities(XTr);
@@ -12,41 +12,30 @@ clusteredDim = [46, 6];
 
 %% Clustering of the data with kmeansStable
 % Need to do that with non-normalised data !
-X = XTr;
-y = yTr;
-[cluster, centers, ~] = kmeansStable([X(:,clusteredDim(1)) y], 2);
-[subcluster, subcenters, ~] = kmeansStable([X(cluster==1,clusteredDim(2)) y(cluster==1)], 2);
 
-finalCluster = cluster;
-finalCluster(cluster == 2) = 3;
-finalCluster(cluster==1) = subcluster;
-finalCenters = [subcenters(1,:);centers(2,:);subcenters(2,:)];
+[clusters, predictClusters] = trainClusters(XTr, yTr, clusteredDim(1), clusteredDim(2));
+numClusters = length(unique(clusters));
+%% Predict test clusters 
 
-%% Predict test clusters using 55 and 6th dimensions
-clusterPred = findClosestGroup([XTe(:,clusteredDim(1))], centers(:,1));
-finalClusterPred = clusterPred;
-finalClusterPred(clusterPred == 2) = 3;
+clustersPred = predictClusters(XTe);
 
-subClusterPred = findClosestGroup([XTe(finalClusterPred~=3,clusteredDim(2))], finalCenters(1:2,1));
-finalClusterPred(finalClusterPred~=3) = subClusterPred;
-
-% figure()
-% plot(XTe(finalClusterPred==1,clusteredDim(1)),XTe(finalClusterPred==1,clusteredDim(2)),'*');
-% hold on
-% plot(XTe(finalClusterPred==2,clusteredDim(1)),XTe(finalClusterPred==2,clusteredDim(2)),'*');
-% plot(XTe(finalClusterPred==3,clusteredDim(1)),XTe(finalClusterPred==3,clusteredDim(2)),'*');
+%figure()
+%plot(XTe(clustersPred == 1, clusteredDim(1)),XTe(clustersPred == 1, clusteredDim(2)),'r*');
+%hold on
+%plot(XTe(clustersPred == 2, clusteredDim(1)),XTe(clustersPred == 2, clusteredDim(2)),'b*');
[…]

### 2015-10-31 src/regression/trainClusters.m

@@ -1,0 +1,29 @@
+function [ finalClusters, predict ] = trainClusters( X, y, dim1, dim2)
+%trainCluster Trains the data to find which cluster a certain point belongs to
+%
+% clusters      clusters for the training data
+% predict       function to predict clusters for the test data
+
+[clusters, centers, ~] = kmeansStable([X(:, dim1) y], 2);
+[subclusters, subcenters, ~] = kmeansStable([X(clusters == 1, dim2) y(clusters == 1)], 2);
+
+finalClusters = clusters;
+finalClusters(clusters == 2) = 3;
+finalClusters(clusters == 1) = subclusters;
+
+predict = @predictCluster;
+
+
+function finalClustersPred = predictCluster(XTe)
+
+clustersPred = findClosestGroup(XTe(:, dim1), centers(:, 1));
+subClustersPred = findClosestGroup(XTe(clustersPred == 1, dim2), subcenters(:, 1));
+
+finalClustersPred = clustersPred;
+finalClustersPred(clustersPred == 2) = 3;
+finalClustersPred(clustersPred == 1) = subClustersPred;
+
+end
+
+end
+

### 2015-10-31 src/utils/findClosestGroup.m

@@ -2,13 +2,6 @@ function [ groups ] = findClosestGroup( X, centers )
 %findClosestGroup Find group of the data by computing the minimum distance
 % to the centers of the groups
 
-for line = 1:length(X)
-    distances = zeros(size(centers,1), 1);
-    for group = 1:size(centers,1)
-        distances(group) = norm(centers(group, :) - X(line, :));
-    end
-    group = find(distances == min(distances));
-    groups(line) = group;
-end
-groups = groups';
+groups = dsearchn(centers, X);
+
 end

### 2015-10-31 src/classification/statistics.m
Generalized choosing of the dimension that classifies data in clusters

@@ -2,7 +2,7 @@
 clear
 clc
 
-load('../data/SaoPaulo_regression.mat');
+load('../data/SaoPaulo_classification.mat');
 len = length(X_train);
 dim = size(X_train, 2);
 

### 2015-10-31 src/regression/predictRegression.m

@@ -1,20 +1,19 @@
 function [ yPred ] = predictRegression( XTr, yTr, XTe, param )
 %predictRegression Prediction function for the regression
 
-%% Find cardinalities of data
+%% Find cardinalities of dimensions
 cards = cardinalities(XTr);
-clusteredDim = [55, 6];
 
-%% Remove dimensions with discrete arguments
+%% Remove categorical dimensions 
 XTr = XTr(:,cards == size(XTr,1));
 XTe = XTe(:,cards == size(XTr,1));
-clusteredDim = [46, 6];
 
-%% Clustering of the data with kmeansStable
+%% Pre clustering of the data
 % Need to do that with non-normalised data !
 
-[clusters, predictClusters] = trainClusters(XTr, yTr, clusteredDim(1), clusteredDim(2));
-numClusters = length(unique(clusters));
+numClusters = 3;
+[clusters, predictClusters] = trainClusters(XTr, yTr);
+
 %% Predict test clusters 
 
 clustersPred = predictClusters(XTe);

### 2015-10-31 src/regression/statistics.m

@@ -2,7 +2,7 @@
 clear
 clc
 
-load('../data/SaoPaulo_classification.mat');
+load('../data/SaoPaulo_regression.mat');
 len = length(X_train);
 dim = size(X_train, 2);
 

### 2015-10-31 src/regression/trainClusters.m

@@ -1,27 +1,41 @@
-function [ finalClusters, predict ] = trainClusters( X, y, dim1, dim2)
+function [ finalClusters, predict ] = trainClusters( X, y )
 %trainCluster Trains the data to find which cluster a certain point belongs to
 %
 % clusters      clusters for the training data
 % predict       function to predict clusters for the test data
 
-[clusters, centers, ~] = kmeansStable([X(:, dim1) y], 2);
-[subclusters, subcenters, ~] = kmeansStable([X(clusters == 1, dim2) y(clusters == 1)], 2);
+% We have three clusters, and two dimensions that correlate highly with y, 
+% so we use them to separate the clusters
+[I, ~] = correlation(X, y);
 
+% Use most correlated dimensions to separate first two clusters
+[clusters, centers, sumd] = kmeansStable([X(:, I(1)) y], 2);
+
+% Use second most correlated dimension to separate sparse cluster in 2
+clusterSizes = [ sum(clusters == 1); sum(clusters == 2) ];
+clusterSpread = sumd ./ clusterSizes;
+sparseCluster = find(clusterSpread == max(clusterSpread));
+subX = X(clusters == sparseCluster, :);
+suby = y(clusters == sparseCluster, :);
+[subclusters, subcenters, ~] = kmeansStable([subX(: , I(2)) suby], 2);
+
+% Relabel clusters for global labeling
 finalClusters = clusters;
-finalClusters(clusters == 2) = 3;
-finalClusters(clusters == 1) = subclusters;
+finalClusters(clusters ~= sparseCluster) = 3;
+finalClusters(clusters == sparseCluster) = subclusters;
 
 predict = @predictCluster;
 
 
 function finalClustersPred = predictCluster(XTe)
-
-clustersPred = findClosestGroup(XTe(:, dim1), centers(:, 1));
-subClustersPred = findClosestGroup(XTe(clustersPred == 1, dim2), subcenters(:, 1));
+% Predict the clusters for test data
+    
+clustersPred = findClosestGroup(XTe(:, I(1)), centers(:, 1));
+subClustersPred = findClosestGroup(XTe(clustersPred == sparseCluster, I(2)), subcenters(:, 1));
 
 finalClustersPred = clustersPred;
-finalClustersPred(clustersPred == 2) = 3;
-finalClustersPred(clustersPred == 1) = subClustersPred;
+finalClustersPred(clustersPred ~= sparseCluster) = 3;
+finalClustersPred(clustersPred == sparseCluster) = subClustersPred;
 
 end
 

### 2015-10-31 src/regression/predictRegression.m
fix outliers

@@ -8,7 +8,7 @@ cards = cardinalities(XTr);
 XTr = XTr(:,cards == size(XTr,1));
 XTe = XTe(:,cards == size(XTr,1));
 
-%% Pre clustering of the data
+%% Pre clustering of the data 
 % Need to do that with non-normalised data !
 
 numClusters = 3;
@@ -73,9 +73,7 @@ for cl = 1:numClusters
     XTeNormalised = (X - ones(size(X,1),1)*meanX(cl,:))./(ones(size(X,1),1)*stdX(cl,:));
     
     % Fix outliers
-    limitDeviation = 3;
-    XTeNormalised(XTeNormalised > limitDeviation) = limitDeviation;
-    XTeNormalised(XTeNormalised < -limitDeviation) = -limitDeviation;
+    XTeNormalised = fixOutliers(XTeNormalised, 3);
     
     % Reduce the dimension given PCA of training set
     XTeNormalised = (VReduced{cl,1}'*(V{cl,1}*XTeNormalised'))';

### 2015-11-01 src/classification/plotDataClassification.m
Add necessary paths to src files

@@ -3,7 +3,7 @@ close all
 clear
 clc
 
-load('../data/SaoPaulo_classification.mat');
+load('../../data/SaoPaulo_classification.mat');
 
 %X_train = fixOutliers(X_train, 1);
 keep = zeros(1, length(X_train));

### 2015-11-01 src/classification/statistics.m

@@ -2,7 +2,10 @@
 clear
 clc
 
-load('../data/SaoPaulo_classification.mat');
+addpath('../utils');
+addpath('../methods');
+
+load('../../data/SaoPaulo_classification.mat');
 len = length(X_train);
 dim = size(X_train, 2);
 

### 2015-11-01 src/regression/plotData.m

@@ -3,7 +3,7 @@ clear
 clc
 close all
 
-load('../data/SaoPaulo_regression.mat');
+load('../../data/SaoPaulo_regression.mat');
 len = length(X_train);
 dim = size(X_train, 2);
 

### 2015-11-01 src/classification/main.m
Update classification files

@@ -1,0 +1,17 @@
+close all
+clear
+clc
+
+addpath('../utils');
+addpath('../methods');
+
+load('../../data/SaoPaulo_classification.mat')
+
+% predict = @(Xtr, Ytr, Xte) predictRegression(Xtr, Ytr, Xte, 10^5);
+% [ trRMSE, teRMSE ] = crossValidation(y_train, X_train, 3, predict);
+
+[ trRMSE, teRMSE, lambda ] = crossValidationParam(X_train, y_train, 3, @predictClassification, 4);
+
+semilogx(lambda, teRMSE, 'r*-')
+hold on
+semilogx(lambda, trRMSE, 'b*-')

### 2015-11-01 src/classification/predictClassification.m

@@ -1,15 +1,79 @@
-function [ groups ] = predictClassification( tX, beta, k )
-%predictClass Return the most probable class for each data
+function [ yPred, pPred ] = predictClassification( XTr, yTr, XTe, param )
+%predictClassification Prediction function for the classification
 
+%% Separe data set in function of the 20th dimension of X
+cluster = XTr(:,20);
+clusterPred = XTe(:,20);
 
-for class = 1:k
+%% Find cardinalities, remove binary dimensions
+cards = cardinalities( XTr );
+XTr = XTr(:,cards > 2);
+XTe = XTe(:,cards > 2);
+
+%% Train classification model for each cluster
+sizeCluster = length(unique(cluster));
+tXTr = cell(sizeCluster,1);
+yTrF = cell(sizeCluster,1);
+beta = cell(sizeCluster,1);
+
+for cl = 1:sizeCluster
+    % Take the data for one cluster
+    X = XTr(cluster == cl,:);
+    y = yTr(cluster == cl,:);
     
-    yPred(:,class) = sigma(tX * beta(:,class));
+    % Transform y to have only 1 and 0 
+    y(y < 0) = 0;
     
+    % Separate discrete and continuous dimensions
+    cardsX(cl,:) = cardinalities( X );
+    disX = X(:,cardsX(cl,:) ~= size(X,1));
+    contX = X(:,cardsX(cl,:) == size(X,1));
+    
+    % Normalise the continuous data
+    [XTrNormalised, meanX(cl,:), stdX(cl,:)] = normalise(contX);
+    
+    % Remove outliers
+    [XTrFiltered, yTrFiltered, linesKept] = removeOutlierLines(XTrNormalised, y, 3, 1);
+    disX = disX(linesKept == 1,:);
+    
+    % Put discrete and continuous dimensions together again
+    XTrKept = [XTrFiltered, disX];
+    
+    % Compute tX and final y
+    tXTr{cl,1} = [ones(length(XTrKept), 1)  XTrKept];
+    yTrF{cl,1} = yTrFiltered;
+    
+    % Compute (penalised) logistic regression
+%     beta{cl,1} = logisticRegression(yTrF{cl,1}, tXTr{cl,1}, 0.001);
+    beta{cl,1} = penLogisticRegression(yTrF{cl,1}, tXTr{cl,1}, 0.001, 0.01);
+
 end
 
[…]

### 2015-11-01 src/methods/logisticRegression.m

@@ -1,7 +1,7 @@
 function beta = logisticRegression( y, tX, alpha )
 
 % Maximum number of iterations
-maxIters = 1000; % TO DEFINE
+maxIters = 15000; % TO DEFINE
 
 % Number of data
 N = size(tX,1);
@@ -24,7 +24,6 @@ for k = 1:maxIters
     % Compute cost of logistic regression
     L = computeCostLogisticReg(y, tX, beta);
     alpha = abs(L*10^(-4));
-    fprintf('%d : %f  %f\n', k,L, alpha);
     
     % Store beta and L
     beta_all(:,k) = beta;
@@ -34,7 +33,7 @@ for k = 1:maxIters
     
     % Look at the convergence
     if k>1
-%         fprintf('%d : %f\n', k, max(abs(beta-beta_all(:,k-1))));
+%         fprintf('%d : %f  %f %f\n', k,L, alpha, max(abs(beta-beta_all(:,k-1))));
         if abs(beta-beta_all(:,k-1))<0.001
             % If the difference between last and present beta is small...
             % break

### 2015-11-01 src/methods/penLogisticRegression.m

@@ -1,7 +1,7 @@
 function beta = penLogisticRegression(y,tX,alpha,lambda)
 
 % Maximum number of iterations
-maxIters = 1000; % TO DEFINE
+maxIters = 5000; % TO DEFINE
 
 % Number of data
 N = size(tX,1);
@@ -25,7 +25,7 @@ for k = 1:maxIters
     % Compute cost of logistic regression
     L = computeCostPenLogisticReg(y, tX, beta, lambda);
     alpha = abs(L*10^(-4));
-    fprintf('%d : %f  %f\n', k,L, alpha);
+    
     
     % Store beta and L
     beta_all(:,k) = beta;
@@ -33,7 +33,7 @@ for k = 1:maxIters
     
     % Look at the convergence
     if k>1
-%         fprintf('%d : %f\n', k, max(abs(beta-beta_all(:,k-1))));
+%         fprintf('%d : %f  %f %f\n', k,L, alpha, max(abs(beta-beta_all(:,k-1))));
         if abs(beta-beta_all(:,k-1))<0.001
             % If the difference between last and present beta is small...
             % break

### 2015-11-01 src/regression/predictRegression.m

@@ -40,13 +40,9 @@ for cl = 1:numClusters
     [XTrNormalised, meanX(cl,:), stdX(cl,:)] = normalise(X);
     [yTrNormalised, meanY(cl,:), stdY(cl,:)] = normalise(y);
     
-    %% Remove outliers
+    % Fix outliers back to 3 standard deviations
     XTrFiltered = fixOutliers(XTrNormalised, 3);
-%     XTrFiltered = XTrNormalised;
     yTrFiltered = yTrNormalised;
-    % [XTrFiltered, yFiltered] = removeOutlierLines(XTrNormalised, [yTrNormalised finalCluster], 3, 1);
-    % finalCluster = yFiltered(:,2);
-    % yTrFiltered = yFiltered(:,1);
     
     % Remove correlated columns using PCA 
     [XTrKept, V{cl,1}, VReduced{cl,1}] = pca(XTrFiltered, 1);
@@ -55,6 +51,7 @@ for cl = 1:numClusters
     tXTr{cl,1} = [ones(length(XTrKept), 1)  XTrKept];
     yTrF{cl,1} = yTrFiltered;
     
+    % Compute regression
 %     beta{cl,1} = leastSquaresGD(yTrF{cl,1}, tXTr{cl,1}, param);
 %     beta{cl,1} = leastSquares(yTrF{cl,1}, tXTr{cl,1});
     beta{cl,1} = ridgeRegression(yTrF{cl,1}, tXTr{cl,1}, param);
@@ -68,7 +65,7 @@ yPred = zeros(size(XTe,1),1);
 for cl = 1:numClusters
     % Take the data from one cluster
     X = XTe(clustersPred == cl,:);
-    %y = yTr(finalClusterPred == cl,:);
+
     % Normalise given mean and std of the training set
     XTeNormalised = (X - ones(size(X,1),1)*meanX(cl,:))./(ones(size(X,1),1)*stdX(cl,:));
     

### 2015-11-01 src/utils/crossValidation.m

@@ -1,12 +1,12 @@
-function [ trRMSE, teRMSE ] = crossValidation( X, y, groups, predict)
+function [ trError, teError ] = crossValidation( X, y, groups, predict)
 %crossValidationRegression Returns RMSE for train and test for this prediciton function
 % X         Data input
 % y         Output
 % groups    Number of groups to form
 % predict   The prediciton function
 
-trRMSE = zeros(1, groups);
-teRMSE = zeros(1, groups);
+trError = zeros(1, groups);
+teError = zeros(1, groups);
 
 for groupNumber = 1:groups
     %% Separe X and y in train and test parts
@@ -16,8 +16,12 @@ for groupNumber = 1:groups
     yTrPred = predict( XTr, yTr, XTr );
     yTePred = predict( XTr, yTr, XTe );
     
-    trRMSE(groupNumber) = computeRMSE(yTr, yTrPred);
-    teRMSE(groupNumber) = computeRMSE(yTe, yTePred);
+    % RMSE
+%     trError(groupNumber) = computeRMSE(yTr, yTrPred);
+%     teError(groupNumber) = computeRMSE(yTe, yTePred);
+    % Loss function
+    trError(groupNumber) = computeLoss(yTr, yTrPred);
+    teError(groupNumber) = computeLoss(yTe, yTePred);
     
 end
 

### 2015-11-01 src/utils/crossValidationParam.m

@@ -1,7 +1,8 @@
 function [ trRMSE, teRMSE, param ] = crossValidationParam( X, y, groups, predictParam, nbIter)
 %crossValidationRegression
 
-param = logspace(-6, 10, 60);
+% param = logspace(-6, 10, 60); % For ridge regression
+param = logspace(-3, 3, 10); % For penalised logistic regression
 lenParam = length(param);
 
 trRMSE = zeros(1, lenParam);

### 2015-11-01 src/utils/removeOutlierLines.m

@@ -1,4 +1,4 @@
-function [XFiltered, yFiltered] = removeOutlierLines(X, y, limitDeviation, limitPerLine)
+function [XFiltered, yFiltered, linesKept] = removeOutlierLines(X, y, limitDeviation, limitPerLine)
 % removeOutlierLines Removes lines that contain too many outliers
 %
 % limitDeviation    The number of standard deviations away a point is to be
@@ -21,6 +21,10 @@ end
 % Remove all line that have 2 or more outliers
 XFiltered = X(counts < limitPerLine, :);
 yFiltered = y(counts < limitPerLine, :);
+
+linesKept = zeros(1, len);
+linesKept(counts < limitPerLine) = 1;
+linesKept(counts >= limitPerLine) = 0;
 
 end
 

### 2015-11-01 src/classification/main.m
Update classification files

@@ -12,6 +12,15 @@ load('../../data/SaoPaulo_classification.mat')
 
 [ trRMSE, teRMSE, lambda ] = crossValidationParam(X_train, y_train, 3, @predictClassification, 4);
 
-semilogx(lambda, teRMSE, 'r*-')
+% semilogx(lambda, teRMSE, 'r*-')
+% hold on
+% semilogx(lambda, trRMSE, 'b*-')
+
+%%
+figure()
+% subplot(211)
+boxTr = boxplot(trRMSE,'plotstyle','compact','colors','b','labels',lambda);
+% subplot(212)
 hold on
-semilogx(lambda, trRMSE, 'b*-')
+boxTe = boxplot(teRMSE,'plotstyle','compact','colors','r');
+ylim([0.05 0.12])

### 2015-11-01 src/utils/crossValidationParam.m

@@ -5,8 +5,8 @@ function [ trRMSE, teRMSE, param ] = crossValidationParam( X, y, groups, predict
 param = logspace(-3, 3, 10); % For penalised logistic regression
 lenParam = length(param);
 
-trRMSE = zeros(1, lenParam);
-teRMSE = zeros(1, lenParam);
+trRMSE = zeros(nbIter, lenParam);
+teRMSE = zeros(nbIter, lenParam);
 
 for paramIdx = 1:lenParam
     
@@ -25,8 +25,8 @@ for paramIdx = 1:lenParam
     trMeanError = trMeanError( abs(normalise(trMeanError)) < 2 );
     teMeanError = teMeanError( abs(normalise(teMeanError)) < 2 );
 
-    trRMSE(paramIdx) = mean(trMeanError);
-    teRMSE(paramIdx) = mean(teMeanError);
+    trRMSE(:,paramIdx) = trMeanError;
+    teRMSE(:,paramIdx) = teMeanError;
     
     fprintf('Finished %d out of %d\n', paramIdx, lenParam);
     

### 2015-11-01 src/methods/logisticRegression.m
Change alpha parameter of logistic regression

@@ -1,10 +1,13 @@
-function beta = logisticRegression( y, tX, alpha )
+function beta = logisticRegression( y, tX, alphaInitial )
 
 % Maximum number of iterations
 maxIters = 15000; % TO DEFINE
 
 % Number of data
 N = size(tX,1);
+
+% alpha initialisation
+alpha = alphaInitial;
 
 % Beta initialisation
 beta = zeros(size(tX,2),1);

### 2015-11-01 src/methods/penLogisticRegression.m

@@ -1,10 +1,13 @@
-function beta = penLogisticRegression(y,tX,alpha,lambda)
+function beta = penLogisticRegression(y,tX,alphaInitial,lambda)
 
 % Maximum number of iterations
 maxIters = 5000; % TO DEFINE
 
 % Number of data
 N = size(tX,1);
+
+% alpha initialisation
+alpha = alphaInitial;
 
 % Beta initialisation
 beta = zeros(size(tX,2),1);

### 2015-11-02 report/paper.tex
Update report and PCA

@@
 
-\section{Data Description}
-Our train-data consists of ...
+\section{Regression}
+\subsection{Data Description}
+Our regression-data contains $N=2800$ train-examples and $N=1200$ test-data.
+The train-data are composed of input and output variables $X$ and $y$.
+Each input $x_n$ has a dimensionality $D=70$ with 60 real-valued variables, 1 binary variable, 5 and 4 categorical variables with respectively 3 and 4 categories.
+The output of our test-data is not observed.
+Our goal is then to predict $y$ for all test-examples and to give an approximation of the test-error.
 
-\section{Data visualization and cleaning}
-We performed basic exploratory data analysis on our data. Figure \ref{fig:boxplotX} ...
+\subsection{Data Visualization and Cleaning}
+To visualise our data, we plot histograms to see the distribution of $y$ and of each dimension of X. The input data is not centred, so that we need to normalise it.
 
-\begin{figure}[!h] % !t
-\center
-\subfigure[Boxplot of real-valued $\mathbf{X}$. Data is not centered and therefore we normalize it.]{\includegraphics[width=2.5in]{figures/boxplotX.pdf} \label{fig:boxplotX}}
-\hfill
-\subfigure[Histogram of $\mathbf{y}$. We can clearly see two outliers.]{\includegraphics[width=2.5in]{figures/histY.pdf} \label{fig:histY}}
-\caption{}
-\end{figure}
+Figure \todo{ADD FIGURE} shows the histogram of the output variable.
+We see that the data are split in three bursts of different sizes : a large amount of the data have small output values and only a small part have the highest values.
+Figure \todo{ADD FIGURE} shows histograms of two interesting dimensions of $X$.
+As all others real-valued variables seem to have sort of Gaussian distributions, we can see that those are split in two bursts.
+By plotting $y$ in function of each of these variables, as shown in Figure \todo{ADD FIGURE}, we can see that $y$ is highly correlated with dimensions $6$ and $55$ of $X$.
+The output-data can thus be separated in three clusters by looking at $x_n(6)$ ans $x_n(55)$ for each data.
 
-\section{Ridge regression}
+We also computed the eigenvalues of the matrix $X^T X$.
+As ten eigenvalues are smaller than $10^{-10}$, we will need to transform $X$ to remove those small eigenvalues or to use methods such as ridge regression to lift them.
+
+%\begin{figure}[!h] % !t
+%\center
+%\subfigure[Boxplot of real-valued $\mathbf{X}$. Data is not centered and therefore we normalize it.]{\includegraphics[width=2.5in]{figures/boxplotX.pdf} \label{fig:boxplotX}}
+%\hfill
+%\subfigure[Histogram of $\mathbf{y}$. We can clearly see two outliers.]{\includegraphics[width=2.5in]{figures/histY.pdf} \label{fig:histY}}
+%\caption{}
+%\end{figure}
+
+\subsection{Clustering and Feature Transformations}
+
+\subsection{Methods}
 We applied least-squares and ridge regression to this dataset...
 
-\section{Feature transformations}
+\subsection{Feature transformations}
 We tried several feature transformations...
+
+\section{Classification}
+\subsection{Data Description}
+\subsection{Data Visualization and Cleaning}
+\subsection{Logistic Regression}
 

### 2015-11-02 src/regression/main.m

@@ -12,6 +12,6 @@ load('../../data/SaoPaulo_regression.mat')
 
 [ trRMSE, teRMSE, lambda ] = crossValidationParam(X_train, y_train, 3, @predictRegression, 4);
 
-semilogx(lambda, teRMSE, 'r*-')
-hold on
-semilogx(lambda, trRMSE, 'b*-')
+% semilogx(lambda, teRMSE, 'r*-')
+% hold on
+% semilogx(lambda, trRMSE, 'b*-')

### 2015-11-02 src/regression/plotData.m

@@ -10,5 +10,6 @@ dim = size(X_train, 2);
 for d = 1:dim
    figure();
    hold on;
-   plot(X_train(:, d), y_train, '*');
+%    plot(X_train(:, d), y_train, '*');
+   histogram(X_train(:,d));
 end

### 2015-11-02 src/utils/pca.m

@@ -1,8 +1,8 @@
 function [ XReduced, V, VReduced ] = pca( X, minEigen)
 % PCA Do PCA to remove the correlated dimensions of tX
 % cov*V = V*lambda
-normX = normalise(X'*X);
-covariance = cov(normX);
+
+covariance = cov(X'*X);
 [V, lambda] = eig(covariance);
 
 [~, lambdaCol] = find(lambda > minEigen);

### 2015-11-02 src/regression/main.m
Plots

@@ -15,3 +15,13 @@ load('../../data/SaoPaulo_regression.mat')
 semilogx(lambda, teRMSE, 'r*-')
 hold on
 semilogx(lambda, trRMSE, 'b*-')
+
+%%
+figure()
+ax(1) = subplot(211);
+boxTr = boxplot(trRMSE,'plotstyle','compact','colors','b','labels',lambda);
+ax(2) = subplot(212);
+hold on
+boxTe = boxplot(teRMSE,'plotstyle','compact','colors','r');
+linkaxes(ax);
+%ylim([0.05 0.12])

### 2015-11-02 src/regression/generatePlots.m
Plots generation

@@ -1,0 +1,129 @@
+close all
+clear
+clc
+
+addpath('../utils');
+addpath('../methods');
+
+load('../../data/SaoPaulo_regression.mat')
+
+len = length(X_train);
+dim = size(X_train, 2);
+
+FONT_SIZE = 16;
+LABEL_FONT_SIZE = 20;
+
+setFontSize = @(gca, gcf, xl, yl) {
+    set(gca, 'FontSize', FONT_SIZE); 
+    set(findall(gcf,'type','text'), 'FontSize', FONT_SIZE); 
+    set(xl, 'FontSize', LABEL_FONT_SIZE); set(yl, 'FontSize', LABEL_FONT_SIZE)
+    };
+
+
+%% Histogram of Y
+figure();
+histogram(y_train);
+xl = xlabel('y values');
+yl = ylabel('Frequency');
+setFontSize(gca, gcf, xl, yl);
+print -dpdf 'plots/y_histogram.pdf';
+close;
+
+%% Dimension 55
+figure();
+plot(X_train(:, 55), y_train, '*');
+xl = xlabel('55th dimension');
+yl = ylabel('y');
+setFontSize(gca, gcf, xl, yl);
+print -dpdf 'plots/dim55.pdf';
+close;
+
+%% Dimension 6
+figure();
+plot(X_train(:, 6), y_train, '*');
+xl = xlabel('6th dimension');
+yl = ylabel('y');
+setFontSize(gca, gcf, xl, yl);
+print -dpdf 'plots/dim6.pdf';
+close;
+
+%% Plot trained clusters
+figure();
+hold on;
+[cluster, predict] = trainClusters(X_train, y_train);
+plot(X_train(cluster == 1, 55), X_train(cluster == 1, 6), 'r*');
+plot(X_train(cluster == 2, 55), X_train(cluster == 2, 6), 'b*');
+plot(X_train(cluster == 3, 55), X_train(cluster == 3, 6), '*', 'MarkerEdgeColor', [0,0.5,0]);
+legend('Cluster 1', 'Cluster 2', 'Cluster 3');
+xl = xlabel('55th dimension');
+yl = ylabel('6th dimension');
+setFontSize(gca, gcf, xl, yl);
+print -dpdf 'plots/clusters_train.pdf';
+close;
+
+%% Plot test clusters phase 1
+
+figure();
+hold on;
+cluster = predict(X_test);
[…]

### 2015-11-02 src/regression/statistics.m

@@ -41,4 +41,21 @@ strongCorrIdx = find(corrs > 2);
 fprintf('Dimensions with strong correlation to y:');
 disp(I(strongCorrIdx));
 
+%% Eigen values
+section('Eigen values of covariance matrix');
+card = cardinalities(X_train);
+X = X_train(:, card == len);
+covariance = cov(X' * X);
+[~, lambda] = eig(covariance);
+eigValues = diag(lambda);
+numSteps = ceil(log10(max(eigValues)) / 2 ) * 2 ;
 
+for x = logspace(0, numSteps, numSteps / 2 + 1)
+    if( x == 1 )
+        fprintf('Eigen values < 1 : %d\n', sum(eigValues < 1))
+    else 
+        lowBound = x/100;
+        count = sum(eigValues >= lowBound & eigValues < x);
+        fprintf('Eigen values >= %.0E && < %.0E : %d\n', lowBound, x, count);
+    end
+end

### 2015-11-02 report/paper.tex
Update report

@@
 \subsection{Data Visualization and Cleaning}
-To visualise our data, we plot histograms to see the distribution of $y$ and of each dimension of X. The input data is not centred, so that we need to normalise it.
+To visualise our data, we plotted histograms to see the distribution of $y$ and of $X$ dimensions. The input data is not centred, so that we need to normalise it.
 
@@
 We also computed the eigenvalues of the matrix $X^T X$.
-As ten eigenvalues are smaller than $10^{-10}$, we will need to transform $X$ to remove those small eigenvalues or to use methods such as ridge regression to lift them.
+As ten eigenvalues are smaller than $10^{-10}$, we will need to transform $X$ to remove dimensions corresponding to those small eigenvalues or to use methods such as ridge regression to lift them.
 
@@
 
-\subsection{Clustering and Feature Transformations}
+\subsection{Clustering}
+We used k-means clustering to separate our data during the training part. The clustering is always done before normalising the data.
+As the data are not equally shared between the three clusters, we first compute a clustering with $k=2$ clusters using $y$ and the most correlated dimension of $X$ to compute the distance between each data-point and centre.
+Then, we keep only the data classified in the sparse cluster to make a second clustering with $k=2$ using $y$ and the second most correlated dimension of $X$.
+
+In order to ensure a reliable clustering, we do not initialise the centres randomly.
+During the initialisation, we sort the data-points by their norm and separate the sorted list in $k$ chunks, where $k$ is the given number of clusters.
+The mean along each dimension is then calculated for each chunk giving the initial centres for the clusters.
+
+During the test part, we follow the same principle.
+The most correlated dimension of $X$ found during the training part is used to find the closest centre and the second most correlated dimension is then used to divide the data classified in the sparse cluster in two sub-clusters.
+
+Figure \todo{ADD FIGURE} shows an example of k-means clustering.
+
+\subsection{Feature transformations}
+After the separation in cluster, each group of train-data is treated separately.
+For each cluster, the dimensions of $X$ and $y$ are normalised to have zero-mean and unitary standard deviation.
+We also implemented a method to fix the outliers and a PCA to improve our predictions.
+
+We defined as outliers the data-points for which some elements of one or more real-valued dimensions are further than 3 standard deviations from the mean.
+As our data were normalised, all non-categorical element higher than 3 were considered as an outlier element.
+To fix the outliers, value $3$ (or $-3$) was assigned to those elements.
+
+The PCA is applied only on the real-valued dimensions.
+It projects our input data in a new coordinate system based on the covariance between dimensions.
+It allows us to reduce the dimensionality of our data in this new space by removing the dimensions with eigenvalues lower than 1.
+The data are then projected back in the initial coordinate system.
+In our case, the PCA reduce our input data by eleven dimensions.
+\todo{ADD FIGURE WITH EIGENVALUES}
+
+After the clustering of test-data, each group is also treated separately.
+The data are first normalised given the means and standard deviations computed during the training.
+The outliers are then fixed and PCA is applied by projecting the data with the eigenvectors computed during the training part.
 
 \subsection{Methods}
-We applied least-squares and ridge regression to this dataset...
-
-\subsection{Feature transformations}
-We tried several feature transformations...
+We applied least-squares and ridge regression to this dataset... (remove non-real-valued dimensions)
 
@@
 \subsection{Data Description}
+Our classification-data contains $N=1500$ train-examples and $N=1500$ test-data.
+The train-data are composed of input and output variables $X$ and $y$.
+Each input $x_n$ has a dimensionality $D=37$ with 32 real-valued variables, 2 binary variables, 3 categorical variables with 5 categories.
+The output is a binary variable equal to $1$ or $-1$.
+The output of our test-data is not observed.
+Our goal is then to predict $y$ for all test-examples and to give an approximation of the test-error.
+
 \subsection{Data Visualization and Cleaning}
+To visualise our data, we plotted histogram of $X$ dimensions. The input data is again not centred, so that we need to normalise it.
+
+Figure \todo{ADD FIGURE} shows the histogram of the \todo{DIMENSION} variable of $X$.
+We see that the distribution is composed of a high peak and a Gaussian distribution.
+We observed that most of the real-valued dimensions have similar distributions and that data-points belonging to the peak or to the Gaussian in one dimensions are contained in the same part of the distribution in other dimensions.
+Moreover, the binary value of the 20th dimension is different for data-points in the peak or in the Gaussian, so that we can use this dimension to separate the input data in two clusters.
+
+We also computed the eigenvalues of the matrix $X^{T}X$. As the smaller eigenvalue is equal to $58.4$, the matrix is full-rank and we do not need to use methods to lift the eigenvalues.
+
+\subsection{Feature Transformations}
+The train-data are first separated in two clusters given the value of the 20th dimension of $X$.
+Each group of data are then treated separately.
+$X$ real-valued variables are first normalised to have zero-mean and unitary standard deviation.
+We also implemented a method to remove the outliers.
+
+We defined as outliers the data-points for which at least one element of a real-valued dimension is further than 3 standard deviations from the mean.
+In this case, the data point is removed from the training set.
+
+After the clustering of the test-data, each group is also treated separately. The real-valued input data are normalised given the means and standard deviations computed during the training.
+
 \subsection{Logistic Regression}

### 2015-11-02 src/regression/generatePlots.m
Update regression plots

@@ -54,7 +54,7 @@ hold on;
 plot(X_train(cluster == 1, 55), X_train(cluster == 1, 6), 'r*');
 plot(X_train(cluster == 2, 55), X_train(cluster == 2, 6), 'b*');
 plot(X_train(cluster == 3, 55), X_train(cluster == 3, 6), '*', 'MarkerEdgeColor', [0,0.5,0]);
-legend('Cluster 1', 'Cluster 2', 'Cluster 3');
+legend('Cluster 1', 'Cluster 2', 'Cluster 3','Location','northwest');
 xl = xlabel('55th dimension');
 yl = ylabel('6th dimension');
 setFontSize(gca, gcf, xl, yl);
@@ -69,7 +69,7 @@ cluster = predict(X_test);
 plot(X_test(cluster == 3, 55), X_test(cluster == 3, 6), 'b*');
 plot(X_test(cluster == 1, 55), X_test(cluster == 1, 6), 'r*');
 plot(X_test(cluster == 2, 55), X_test(cluster == 2, 6), 'r*');
-legend('Cluster 1', 'Cluster 2');
+legend('Cluster 1', 'Cluster 2','Location','northwest');
 xl = xlabel('55th dimension');
 yl = ylabel('6th dimension');
 setFontSize(gca, gcf, xl, yl);
@@ -84,7 +84,7 @@ cluster = predict(X_test);
 plot(X_test(cluster == 1, 55), X_test(cluster == 1, 6), '*', 'MarkerEdgeColor', [0,0,0]);
 plot(X_test(cluster == 2, 55), X_test(cluster == 2, 6), '*', 'MarkerEdgeColor', [1,0,1]);
 plot(X_test(cluster == 3, 55), X_test(cluster == 3, 6), 'w*');
-legend('SubCluster 1', 'SubCluster 2');
+legend('SubCluster 1', 'SubCluster 2','Location','northwest');
 xl = xlabel('55th dimension');
 yl = ylabel('6th dimension');
 setFontSize(gca, gcf, xl, yl);
@@ -107,8 +107,8 @@ grid on
 grid minor
 hold on
 semilogx(lambda, mean(trRMSE), '*b');
-xl = xlabel('Train RMSE');
-yl = ylabel('lambda');
+xl = xlabel('Lambda');
+yl = ylabel('Train RMSE');
 ylim([450 800])
 xlim([10^-3 10^7])
[…]

### 2015-11-02 src/regression/generatePlots.m
Update plots for regression

@@ -26,7 +26,7 @@ histogram(y_train);
 xl = xlabel('y values');
 yl = ylabel('Frequency');
 setFontSize(gca, gcf, xl, yl);
-print -dpdf 'plots/y_histogram.pdf';
+print -dpng 'plots/y_histogram.png';
 close;
 
 %% Dimension 55
@@ -35,7 +35,7 @@ plot(X_train(:, 55), y_train, '*');
 xl = xlabel('55th dimension');
 yl = ylabel('y');
 setFontSize(gca, gcf, xl, yl);
-print -dpdf 'plots/dim55.pdf';
+print -dpng 'plots/dim55.png';
 close;
 
 %% Dimension 6
@@ -44,7 +44,7 @@ plot(X_train(:, 6), y_train, '*');
 xl = xlabel('6th dimension');
 yl = ylabel('y');
 setFontSize(gca, gcf, xl, yl);
-print -dpdf 'plots/dim6.pdf';
+print -dpng 'plots/dim6.png';
 close;
 
 %% Plot trained clusters
@@ -58,7 +58,7 @@ legend('Cluster 1', 'Cluster 2', 'Cluster 3','Location','northwest');
 xl = xlabel('55th dimension');
 yl = ylabel('6th dimension');
 setFontSize(gca, gcf, xl, yl);
-print -dpdf 'plots/clusters_train.pdf';
+print -dpng 'plots/clusters_train.png';
 close;
 
 %% Plot test clusters phase 1
@@ -73,7 +73,7 @@ legend('Cluster 1', 'Cluster 2','Location','northwest');
 xl = xlabel('55th dimension');
 yl = ylabel('6th dimension');
 setFontSize(gca, gcf, xl, yl);
-print -dpdf 'plots/clusters_phase1.pdf';
+print -dpng 'plots/clusters_phase1.png';
 close;
 
 %% Plot test clusters phase 2
@@ -88,7 +88,7 @@ legend('SubCluster 1', 'SubCluster 2','Location','northwest');
 xl = xlabel('55th dimension');
 yl = ylabel('6th dimension');
 setFontSize(gca, gcf, xl, yl);
-print -dpdf 'plots/clusters_phase2.pdf';
+print -dpng 'plots/clusters_phase2.png';
 close;
 
 %% Plot lambda vs RMSE for ridge regression
@@ -124,6 +124,6 @@ yl = ylabel('Test RMSE');
[…]

### 2015-11-02 src/classification/generatePlots.m
Add plots for classification

@@ -1,0 +1,133 @@
+close all
+clear
+clc
+
+addpath('../utils');
+addpath('../methods');
+
+load('../../data/SaoPaulo_classification.mat')
+
+len = length(X_train);
+dim = size(X_train, 2);
+
+FONT_SIZE = 16;
+LABEL_FONT_SIZE = 20;
+
+setFontSize = @(gca, gcf, xl, yl) {
+    set(gca, 'FontSize', FONT_SIZE); 
+    set(findall(gcf,'type','text'), 'FontSize', FONT_SIZE); 
+    set(xl, 'FontSize', LABEL_FONT_SIZE); set(yl, 'FontSize', LABEL_FONT_SIZE)
+    };
+
+
+%% Histogram of dimension 35
+figure();
+X35 = X_train(:, 35);
+
+[~, xbins] = hist(X35, length(unique(floor(X35))));
+type1 = X35(X_train(:, 20) == 1);
+type2 = X35(X_train(:, 20) == 2);
+bars1 = hist(type1, xbins);
+bars2 = hist(type2, xbins);
+b = bar(xbins, [bars1 ; bars2]', 'stacked');
+b(1).FaceColor = 'b';
+b(2).FaceColor = 'r';
+l = cell(1,2);
+l{1}='20th dimension == 1'; 
+l{2}='20th dimension == 2';    
+legend(b,l);
+xl = xlabel('35th dimension''s values');
+yl = ylabel('Frequency');
+setFontSize(gca, gcf, xl, yl);
+
+print -dpng 'plots/dim35.png';
+close;
+
+%% Histogram of dimension 3
+figure();
+X23 = X_train(:, 23);
+
+[~, xbins] = hist(X23, length(unique(round(X23, -2))));
+type1 = X23(X_train(:, 20) == 1);
+type2 = X23(X_train(:, 20) == 2);
+bars1 = hist(type1, xbins);
+bars2 = hist(type2, xbins);
+b = bar(xbins, [bars1 ; bars2]', 'stacked');
+b(1).FaceColor = 'b';
+b(2).FaceColor = 'r';
+l = cell(1,2);
+l{1}='20th dimension == 1'; 
+l{2}='20th dimension == 2';    
+legend(b,l);
+xl = xlabel('3rd dimension''s values');
+yl = ylabel('Frequency');
+setFontSize(gca, gcf, xl, yl);
+
+print -dpng 'plots/dim23.png';
+close;
+
+% Dimension 55
+% figure();
[…]

### 2015-11-02 src/classification/generatePlots.m
remove useless comment

@@ -66,68 +66,3 @@ setFontSize(gca, gcf, xl, yl);
 print -dpng 'plots/dim23.png';
 close;
 
-% Dimension 55
-% figure();
-% plot(X_train(:, 55), y_train, '*');
-% xl = xlabel('55th dimension');
-% yl = ylabel('y');
-% setFontSize(gca, gcf, xl, yl);
-% print -dpdf 'dim55.pdf';
-% close;
-% 
-% Dimension 6
-% figure();
-% plot(X_train(:, 6), y_train, '*');
-% xl = xlabel('6th dimension');
-% yl = ylabel('y');
-% setFontSize(gca, gcf, xl, yl);
-% print -dpdf 'dim6.pdf';
-% close;
-% 
-% Plot trained clusters
-% figure();
-% hold on;
-% [cluster, predict] = trainClusters(X_train, y_train);
-% plot(X_train(cluster == 1, 55), X_train(cluster == 1, 6), 'r*');
-% plot(X_train(cluster == 2, 55), X_train(cluster == 2, 6), 'b*');
-% plot(X_train(cluster == 3, 55), X_train(cluster == 3, 6), '*', 'MarkerEdgeColor', [0,0.5,0]);
-% legend('Cluster 1', 'Cluster 2', 'Cluster 3');
-% xl = xlabel('55th dimension');
-% yl = ylabel('6th dimension');
-% setFontSize(gca, gcf, xl, yl);
-% print -dpdf 'clusters_train.pdf';
-% close;
-% 
-% Plot test clusters phase 1
-% 
-% figure();
-% hold on;
-% cluster = predict(X_test);
-% plot(X_test(cluster == 3, 55), X_test(cluster == 3, 6), 'b*');
-% plot(X_test(cluster == 1, 55), X_test(cluster == 1, 6), 'r*');
-% plot(X_test(cluster == 2, 55), X_test(cluster == 2, 6), 'r*');
-% legend('Cluster 1', 'Cluster 2');
-% xl = xlabel('55th dimension');
-% yl = ylabel('6th dimension');
-% setFontSize(gca, gcf, xl, yl);
-% print -dpdf 'clusters_phase1.pdf';
-% close;
-% 
-% Plot test clusters phase 2
-% 
-% figure();
-% hold on;
-% cluster = predict(X_test);
[…]

### 2015-11-02 report/paper.tex
Update the report and the computeLogLoss function

@@
 
-Figure \todo{ADD FIGURE} shows the histogram of the output variable.
+Figure ~\ref{fig:reg_yHistogram} shows the histogram of the output variable.
 We see that the data are split in three bursts of different sizes : a large amount of the data have small output values and only a small part have the highest values.
-Figure \todo{ADD FIGURE} shows histograms of two interesting dimensions of $X$.
-As all others real-valued variables seem to have sort of Gaussian distributions, we can see that those are split in two bursts.
-By plotting $y$ in function of each of these variables, as shown in Figure \todo{ADD FIGURE}, we can see that $y$ is highly correlated with dimensions $6$ and $55$ of $X$.
+All real-valued variables seem to have sort of Gaussian distributions, excepted $6^{th}$ and $55^{th}$ dimensions that are split in two bursts.
+By plotting $y$ in function of each of these variables, as shown in Figures ~\ref{fig:reg_dim6} and ~\ref{fig:reg_dim55}, we can see that $y$ is highly correlated with dimensions $6$ and $55$ of $X$.
 The output-data can thus be separated in three clusters by looking at $x_n(6)$ ans $x_n(55)$ for each data.
+
+\begin{figure}[!h] % !t
+	\center
+	\includegraphics[width=2.6in]{figures/y_histogram.png}
+	\caption{Histogram of $y$. We can see three bursts of data.}
+	\label{fig:reg_yHistogram}
+\end{figure}
 
@@
 
-%\begin{figure}[!h] % !t
-%\center
-%\subfigure[Boxplot of real-valued $\mathbf{X}$. Data is not centered and therefore we normalize it.]{\includegraphics[width=2.5in]{figures/boxplotX.pdf} \label{fig:boxplotX}}
-%\hfill
-%\subfigure[Histogram of $\mathbf{y}$. We can clearly see two outliers.]{\includegraphics[width=2.5in]{figures/histY.pdf} \label{fig:histY}}
-%\caption{}
-%\end{figure}
+\begin{figure}[!h] % !t
+\center
+\subfigure[$y$ in function of $55^{th}$ dimension of $X$.]{\includegraphics[width=2.6in]{figures/dim6.png} \label{fig:reg_dim6}}
+\hfill
+\subfigure[$y$ in function of $6^{th}$ dimension of $X$.]{\includegraphics[width=2.6in]{figures/dim55.png} \label{fig:reg_dim55}}
+\caption{Graphs of $y$ in function of the input dimensions most correlated with the output.}
+\end{figure}
 
@@
 We used k-means clustering to separate our data during the training part. The clustering is always done before normalising the data.
-As the data are not equally shared between the three clusters, we first compute a clustering with $k=2$ clusters using $y$ and the most correlated dimension of $X$ to compute the distance between each data-point and centre.
+As the data are not equally shared between the three clusters, we first compute a clustering with $k=2$ clusters using $y$ and the most correlated dimension of $X$ with $y$ to compute the distance between each data-point and centre.
 Then, we keep only the data classified in the sparse cluster to make a second clustering with $k=2$ using $y$ and the second most correlated dimension of $X$.
+Figure ~\ref{fig:reg_clustersTrain} shows an example of clusters obtained after the training.
+
+\begin{figure}[!h] % !t
+	\center
+	\includegraphics[width=2.6in]{figures/clusters_train.png}
+	\caption{Clusters obtained along the two most correlated dimensions. We can see three clearly delimited groups of data.}
+	\label{fig:reg_clustersTrain}
+\end{figure}
 
@@
 
-Figure \todo{ADD FIGURE} shows an example of k-means clustering.
+Figure ~\ref{fig:reg_clustersTest} shows an example of k-means clustering for test data.
+
+\begin{figure}[!h] % !t
+	\center
+	\subfigure[First k-means applied along the most correlated dimension of $X$ with $y$.]{\includegraphics[width=2.6in]{figures/clusters_phase1.png} \label{fig:reg_clustersTest1}}
+	\hfill
+	\subfigure[Second k-means applied along the second most correlated dimension of $X$ with $y$.]{\includegraphics[width=2.6in]{figures/clusters_phase2.png} \label{fig:reg_clustersTest2}}
+	\caption{Steps of k-means clustering for test-data.}
+	\label{fig:reg_clustersTest}
+\end{figure}
 
@@
 In our case, the PCA reduce our input data by eleven dimensions.
-\todo{ADD FIGURE WITH EIGENVALUES}
 
@@
 
-Figure \todo{ADD FIGURE} shows the histogram of the \todo{DIMENSION} variable of $X$.
+Figures ~\ref{fig:class_dim23} and ~\ref{fig:class_dim35} show respectively the histogram of the $23^{th}$ and the $35^{th}$ variable of $X$.
 We see that the distribution is composed of a high peak and a Gaussian distribution.
 We observed that most of the real-valued dimensions have similar distributions and that data-points belonging to the peak or to the Gaussian in one dimensions are contained in the same part of the distribution in other dimensions.
-Moreover, the binary value of the 20th dimension is different for data-points in the peak or in the Gaussian, so that we can use this dimension to separate the input data in two clusters.
+Moreover, the binary value of the $20^{th}$ dimension is different for data-points in the peak or in the Gaussian, so that we can use this dimension to separate the input data in two clusters.
 
 We also computed the eigenvalues of the matrix $X^{T}X$. As the smaller eigenvalue is equal to $58.4$, the matrix is full-rank and we do not need to use methods to lift the eigenvalues.
+
+\begin{figure}[!h] % !t
+	\center
+	\subfigure[Histogram of the $23^{th}$ dimension of $X$.]{\includegraphics[width=2.6in]{figures/dim23.png} \label{fig:class_dim23}}
+	\hfill
+	\subfigure[Histogram of the $35^{th}$ dimension of $X$.]{\includegraphics[width=2.6in]{figures/dim35.png} \label{fig:class_dim35}}
+	\caption{Example of histogram to show the distribution of $X$ dimensions.
+We can see that each distribution is composed of a peak in blue and of a Gaussian in red, separable given the $20^{th}$ variable.}
+\end{figure}
 

### 2015-11-02 src/regression/main.m
Rafactor predict regression for improved genericity

@@ -10,8 +10,7 @@ load('../../data/SaoPaulo_regression.mat')
 % predict = @(Xtr, Ytr, Xte) predictRegression(Xtr, Ytr, Xte, 10^5);
 % [ trRMSE, teRMSE ] = crossValidation(y_train, X_train, 3, predict);
 
-[ trRMSE, teRMSE, lambda ] = crossValidationParam(X_train, y_train, 3, @predictRegression, 4);
-
+[ trRMSE, teRMSE, lambda ] = findBestRidgeLambda(X_train, y_train);
 %%
 figure()
 ax(1) = subplot(211);

### 2015-11-02 src/regression/predictRegression.m

@@ -1,4 +1,4 @@
-function [ yPred ] = predictRegression( XTr, yTr, XTe, param )
+function [ yPred ] = predictRegression( XTr, yTr, XTe, localPredictor )
 %predictRegression Prediction function for the regression
 
 %% Find cardinalities of dimensions
@@ -54,8 +54,10 @@ for cl = 1:numClusters
     % Compute regression
 %     beta{cl,1} = leastSquaresGD(yTrF{cl,1}, tXTr{cl,1}, param);
 %     beta{cl,1} = leastSquares(yTrF{cl,1}, tXTr{cl,1});
-    beta{cl,1} = ridgeRegression(yTrF{cl,1}, tXTr{cl,1}, param);
-    
+ %   beta{cl,1} = ridgeRegression(yTrF{cl,1}, tXTr{cl,1}, param);
+ 
+    beta{cl, 1} = localPredictor(yTrF{cl,1}, tXTr{cl,1});
+ 
 %     figure()
 %     plot(XTrKept(:,1),yTrF{cl,1},'*');
 end

### 2015-11-02 src/utils/crossValidation.m

@@ -1,12 +1,20 @@
-function [ trError, teError ] = crossValidation( X, y, groups, predict)
+function [ trError, teError ] = crossValidation( X, y, groups, predict, computeLoss)
 %crossValidationRegression Returns RMSE for train and test for this prediciton function
-% X         Data input
-% y         Output
-% groups    Number of groups to form
-% predict   The prediciton function
+% X             Data input
+% y             Output
+% groups        Number of groups to form
+% predict       The prediciton function
+% computeLoss   Set to true to compute loss
 
-trError = zeros(1, groups);
-teError = zeros(1, groups);
+trRMSE = zeros(1, groups);
+teRMSE = zeros(1, groups);
+
+if computeLoss
+    trLoss = zeros(1, groups);
+    teLoss = zeros(1, groups);
+    trLogLoss = zeros(1, groups);
+    teLogLoss = zeros(1, groups);
+end
 
 for groupNumber = 1:groups
     %% Separe X and y in train and test parts
@@ -16,13 +24,38 @@ for groupNumber = 1:groups
     yTrPred = predict( XTr, yTr, XTr );
     yTePred = predict( XTr, yTr, XTe );
     
-    % RMSE
-%     trError(groupNumber) = computeRMSE(yTr, yTrPred);
-%     teError(groupNumber) = computeRMSE(yTe, yTePred);
-    % Loss function
-    trError(groupNumber) = computeLoss(yTr, yTrPred);
-    teError(groupNumber) = computeLoss(yTe, yTePred);
+    % TODO HACK FIXME: for some reason we get cells instea of arrays
+    if iscell(yTrPred) 
+        yTrPred = yTrPred{1};
+    end
+    if iscell(yTePred) 
+        yTePred = yTePred{1};
+    end
     
+    
+    % RMSE
+    trRMSE(groupNumber) = computeRMSE(yTr, yTrPred);
+    teRMSE(groupNumber) = computeRMSE(yTe, yTePred);
+    
+    % Loss function
+    if computeLoss
+        trLoss(groupNumber) = computeLoss(yTr, yTrPred);
+        teLoss(groupNumber) = computeLoss(yTe, yTePred);
+        
+        trLogLoss(groupNumber) = computeLogLoss(yTr, yTrPred);
+        teLogLoss(groupNumber) = computeLogLoss(yTe, yTePred);
+    end
+end
+
+trError.RMSE = trRMSE;
[…]

### 2015-11-02 src/utils/crossValidationParam.m

@@ -1,36 +1,0 @@
-function [ trRMSE, teRMSE, param ] = crossValidationParam( X, y, groups, predictParam, nbIter)
-%crossValidationRegression
-
-% param = logspace(-6, 10, 60); % For ridge regression
-param = logspace(-3, 3, 10); % For penalised logistic regression
-lenParam = length(param);
-
-trRMSE = zeros(nbIter, lenParam);
-teRMSE = zeros(nbIter, lenParam);
-
-for paramIdx = 1:lenParam
-    
-    lambda = param(paramIdx);
-    predict = @(Xtr, ytr, Xte) predictParam(Xtr, ytr, Xte, lambda);
-    
-    trMeanError = zeros(1, nbIter);
-    teMeanError = zeros(1, nbIter);
-    
-    for iter = 1:nbIter
-        [trError, teError] = crossValidation(X, y, groups, predict);
-        trMeanError(iter) = mean(trError);
-        teMeanError(iter) = mean(teError);
-    end
-    
-    trMeanError = trMeanError( abs(normalise(trMeanError)) < 2 );
-    teMeanError = teMeanError( abs(normalise(teMeanError)) < 2 );
-
-    trRMSE(:,paramIdx) = trMeanError;
-    teRMSE(:,paramIdx) = teMeanError;
-    
-    fprintf('Finished %d out of %d\n', paramIdx, lenParam);
-    
-end
-
-end
-

### 2015-11-02 report/paper.tex
Update report

@@
 \begin{abstract}
-In this report, ...
+In this report, we 
 \end{abstract}
@@
 As the data are not equally shared between the three clusters, we first compute a clustering with $k=2$ clusters using $y$ and the most correlated dimension of $X$ with $y$ to compute the distance between each data-point and centre.
-Then, we keep only the data classified in the sparse cluster to make a second clustering with $k=2$ using $y$ and the second most correlated dimension of $X$.
+In a second step, we keep only the data classified in the sparse cluster to make a second clustering with $k=2$ using $y$ and the second most correlated dimension of $X$.
 Figure ~\ref{fig:reg_clustersTrain} shows an example of clusters obtained after the training.
@@
 We defined as outliers the data-points for which some elements of one or more real-valued dimensions are further than 3 standard deviations from the mean.
-As our data were normalised, all non-categorical element higher than 3 were considered as an outlier element.
+As our data were normalised, all non-categorical element higher than 3 were considered as outlier elements.
 To fix the outliers, value $3$ (or $-3$) was assigned to those elements.
+Fixing the outliers with this methods allows us to avoid errors in prediction models due to false-classified data.
 
@@
 It projects our input data in a new coordinate system based on the covariance between dimensions.
-It allows us to reduce the dimensionality of our data in this new space by removing the dimensions with eigenvalues lower than 1.
+It allows us to reduce the dimensionality of our data in this new space by removing the dimensions with eigenvalues lower than a defined value.
+We choose this value equal to $1$ as eleven dimensions have eigenvalues in the order of $10^{-10}$ and all others are higher than $10$.
 The data are then projected back in the initial coordinate system.
@@
 
-\subsection{Methods}
-We applied least-squares and ridge regression to this dataset... (remove non-real-valued dimensions)
+\subsection{Least-Squares and Ridge Regression}
+We applied least-squares, least-squares with gradient descent and ridge regression to our dataset.
+Using PCA allows us to solve the problem of ill-conditioned matrix, so that we can use the three methods mentioned above.
+The ridge regression is the only method that can be used without reducing the dimensionality of the input data.
+
+We computed the RMSE for each model with and without the categorical dimensions of X.
+As removing those dimensions does not change significantly the error, we decided to compute the regression only with the 60 real-valued dimensions.
+
+Figure ~\ref{fig:reg_alphaRMSE} shows the train and test RMSE for a step size $\alpha$ varied from $10^{-3}$ to $10^{-0.5}$ with $20$ points in between.
+The convergence condition was reached when the variation of $\beta$ between two iterations was lower than $1.5 \alpha$.
+Cross-validation was made using $66 \% $ of the data as train data and the rest as test data.
+
+\begin{figure}[!h] % !t
+	\center
+	\includegraphics[width=4in]{figures/lambdaRMSE.png}
+	\caption{Ridge regression for a $66 \%$ - $33 \%$ split with 3 iterations for each $\lambda$.}
+	\label{fig:reg_alphaRMSE}
+\end{figure}
+
+Figure ~\ref{fig:reg_lambdaRMSE} shows the train and test RMSE for $\lambda$ varied from $10^{-3}$ to $10^{7}$ with $60$ points in between.
+Cross-validation was made using $66 \% $ of the data as train data and the rest as test data.
+We can see that $\lambda$ should be chosen between $0$ and $100$ to have the best prediction and that $\lambda$ higher than $1000$ increases the RMSE and the model under-fits.
+
+\begin{figure}[!h] % !t
+	\center
+	\includegraphics[width=4in]{figures/lambdaRMSE.png}
+	\caption{Ridge regression for a $66 \%$ - $33 \%$ split with 3 iterations for each $\lambda$.}
+	\label{fig:reg_lambdaRMSE}
+\end{figure}
+
+Figure TODO compares the three regression methods with $\alpha = TODO$ for the gradient descent and $\lambda = TODO$ for ridge regression. We see that TODO.
 
@@
 Each group of data are then treated separately.
-$X$ real-valued variables are first normalised to have zero-mean and unitary standard deviation.
+The real-valued variables of $X$ are first normalised to have zero-mean and unitary standard deviation.
 We also implemented a method to remove the outliers.
@@
 \subsection{Logistic Regression}
+We applied logistic regression and penalised logistic regression on our dataset.
+Both regressions were implemented with Newton's method.
+To improve the convergence, the step size was adapted at each time step to be proportional to the log likelihood such that $\alpha = log(\ell) * 10^{-4}$.
+The convergence criterion was reached when the difference between $\beta$ between two consecutive iterations was lower than $0.001$.
 

### 2015-11-02 src/methods/leastSquaresGD.m
Plots

@@ -2,7 +2,7 @@ function beta = leastSquaresGD( y, tX, alpha )
 
 % Maximum number of iterations
 maxIters = 10000; % TO DEFINE
-epsilon = 0.001;
+epsilon = alpha * 1.5;
 
 % Number of data
 N = size(tX,1);

### 2015-11-02 src/regression/generatePlots.m

@@ -16,7 +16,17 @@ LABEL_FONT_SIZE = 20;
 setFontSize = @(gca, gcf, xl, yl) {
     set(gca, 'FontSize', FONT_SIZE); 
     set(findall(gcf,'type','text'), 'FontSize', FONT_SIZE); 
-    set(xl, 'FontSize', LABEL_FONT_SIZE); set(yl, 'FontSize', LABEL_FONT_SIZE)
+    set(xl, 'FontSize', LABEL_FONT_SIZE); set(yl, 'FontSize', LABEL_FONT_SIZE);
+    };
+
+setFontSizeNoTouchTicks = @(gca, gcf, xl, yl) {
+    %set(gca, 'FontSize', FONT_SIZE); 
+    %set(findall(gcf,'type','text'), 'FontSize', FONT_SIZE); 
+    set(xl, 'FontSize', LABEL_FONT_SIZE); set(yl, 'FontSize', LABEL_FONT_SIZE);
+    };
+
+getShortLabels = @(nums) {
+    arrayfun(@(x) sprintf('%.1E', x), nums, 'UniformOutput', false);
     };
 
 
@@ -93,20 +103,20 @@ close;
 
 %% Plot lambda vs RMSE for ridge regression
 
-[ trRMSE, teRMSE, lambda ] = crossValidationParam(X_train, y_train, 3, @predictRegression, 3);
+[ trRMSE, teRMSE, lambda ] = findBestRidgeLambda(X_train, y_train);
 
 %%
 figure()
-newLambda = logspace(-3, 7, 300);
+newLambda = logspace(log10(min(lambda)), log10(max(lambda)), 300);
 newTrRMSE = spline(lambda, mean(trRMSE), newLambda);
 newTeRMSE = spline(lambda, mean(teRMSE), newLambda);
 
 ax(1) = subplot(211);
-boxTr = semilogx(newLambda, newTrRMSE, '-b');
+semilogx(newLambda, newTrRMSE, '-b');
 grid on
 grid minor
 hold on
-semilogx(lambda, mean(trRMSE), '*b');
+semilogx(lambda, mean(trRMSE), '.b', 'MarkerSize', 20);
 xl = xlabel('Lambda');
 yl = ylabel('Train RMSE');
 ylim([450 800])
@@ -114,11 +124,11 @@ xlim([10^-3 10^7])
 setFontSize(gca, gcf, xl, yl);
 
 ax(2) = subplot(212);
[…]

### 2015-11-02 report/paper.tex
Refactored part of the first sections
Update report and figures

@@
 \begin{abstract}
-In this report, we 
+In this report, we study several linear method to compute regression and classification on two different datasets.
+For the regression part, we used k-means clustering to separate the data in three clusters.
+As the input matrix is ill-conditioned, we applied a PCA algorithm to reduce its dimensionality.
+WHICH ALGORITHM IS THE BEST; IDEA OF ERROR.
+For the classification part, we separate the data given their distribution along the dimensions of the input.
+WHICH ALGORITHM IS THE BEST; IDEA OF ERROR.
 \end{abstract}
@@
 \subsection{Data Description}
-Our regression-data contains $N=2800$ train-examples and $N=1200$ test-data.
-The train-data are composed of input and output variables $X$ and $y$.
-Each input $x_n$ has a dimensionality $D=70$ with 60 real-valued variables, 1 binary variable, 5 and 4 categorical variables with respectively 3 and 4 categories.
-The output of our test-data is not observed.
-Our goal is then to predict $y$ for all test-examples and to give an approximation of the test-error.
+Our regression-data contains $N=2800$ train-examples and $N=1200$ test-data. The
+train-examples are composed of input and output variables $X$ and $y$. Each
+input $x_n$ has a dimensionality $D=70$ with 60 real-valued variables, 1 binary
+variable, 5 and 4 categorical variables with respectively 3 and 4 categories.
+The output of our test-data is not observed. Our goal is then to predict $y$ for
+all test-examples and to give an approximation of the test-error.
 
 \subsection{Data Visualization and Cleaning}
-To visualise our data, we plotted histograms to see the distribution of $y$ and of $X$ dimensions. The input data is not centred, so that we need to normalise it.
+To visualise our data, we plotted histograms to see the distribution of $y$ and
+of $X$ dimensions. The input data is not centred, so that we need to normalise
+it.
 
-Figure ~\ref{fig:reg_yHistogram} shows the histogram of the output variable.
-We see that the data are split in three bursts of different sizes : a large amount of the data have small output values and only a small part have the highest values.
-All real-valued variables seem to have sort of Gaussian distributions, excepted $6^{th}$ and $55^{th}$ dimensions that are split in two bursts.
-By plotting $y$ in function of each of these variables, as shown in Figures ~\ref{fig:reg_dim6} and ~\ref{fig:reg_dim55}, we can see that $y$ is highly correlated with dimensions $6$ and $55$ of $X$.
-The output-data can thus be separated in three clusters by looking at $x_n(6)$ ans $x_n(55)$ for each data.
+Figure ~\ref{fig:reg_yHistogram} shows the histogram of the output variable. We
+see that the data are split in three bursts of different sizes : a large amount
+of the data have small output values and only a small part have the highest
+values. All real-valued variables seem to have sort of Gaussian distributions,
+excepted $6^{th}$ and $55^{th}$ dimensions that are split in two bursts. By
+plotting $y$ in function of each of these variables, as shown in Figures
+~\ref{fig:reg_dim6} and ~\ref{fig:reg_dim55}, we can see that $y$ is highly
+correlated with dimensions $6$ and $55$ of $X$. The output-data can thus be
+separated in three clusters by looking at $x_n(6)$ ans $x_n(55)$ for each data.
 
@@
 
+Figure TODO shows the train and test TODO ERROR for $\lambda$ varied from
+
 \section{Summary}

### 2015-11-02 src/classification/main.m
classification refactoring

@@ -10,7 +10,7 @@ load('../../data/SaoPaulo_classification.mat')
 % predict = @(Xtr, Ytr, Xte) predictRegression(Xtr, Ytr, Xte, 10^5);
 % [ trRMSE, teRMSE ] = crossValidation(y_train, X_train, 3, predict);
 
-[ trRMSE, teRMSE, lambda ] = crossValidationParam(X_train, y_train, 3, @predictClassification, 4);
+[ trRMSE, teRMSE, trLoss, teLoss, lambda ] = findBestPenLambda(X_train, y_train);
 
 % semilogx(lambda, teRMSE, 'r*-')
 % hold on

### 2015-11-02 src/classification/runClassificationParam.m

@@ -1,0 +1,46 @@
+function [ trRMSE, teRMSE, trLoss, teLoss ] = runClassificationParam( X, y, predictParam, params )
+
+groups = 2; % Number of groups
+nbIter = 3; % Number of times to run cross validation
+
+lenParam = length(params); 
+
+trRMSE = zeros(nbIter, lenParam);
+teRMSE = zeros(nbIter, lenParam);
+
+trLoss = zeros(nbIter, lenParam);
+teLoss = zeros(nbIter, lenParam);
+
+for paramIdx = 1:lenParam
+    
+    lambda = params(paramIdx);
+    predict = @(Xtr, ytr, Xte) predictParam(Xtr, ytr, Xte, lambda);
+    
+    trMeanRMSE = zeros(1, nbIter);
+    teMeanRMSE = zeros(1, nbIter);
+    
+    trMeanLoss = zeros(1, nbIter);
+    teMeanLoss = zeros(1, nbIter);
+    
+    for iter = 1:nbIter
+        [trError, teError] = crossValidation(X, y, groups, predict, true);
+        
+        trMeanRMSE(iter) = mean(trError.RMSE);
+        teMeanRMSE(iter) = mean(teError.RMSE);
+        
+        trMeanLoss(iter) = mean(trError.Loss);
+        teMeanLoss(iter) = mean(teError.Loss);
+    end
+    
+    trRMSE(:,paramIdx) = trMeanRMSE;
+    teRMSE(:,paramIdx) = teMeanRMSE;
+    
+    trLoss(:,paramIdx) = trMeanLoss;
+    teLoss(:,paramIdx) = teMeanLoss;
+    
+    fprintf('Finished %d out of %d\n', paramIdx, lenParam);
+    
+end
+
+end
+

### 2015-11-02 src/utils/crossValidation.m

@@ -1,15 +1,15 @@
-function [ trError, teError ] = crossValidation( X, y, groups, predict, computeLoss)
+function [ trError, teError ] = crossValidation( X, y, groups, predict, doComputeLoss)
 %crossValidationRegression Returns RMSE for train and test for this prediciton function
-% X             Data input
-% y             Output
-% groups        Number of groups to form
-% predict       The prediciton function
-% computeLoss   Set to true to compute loss
+% X                 Data input
+% y                 Output
+% groups            Number of groups to form
+% predict           The prediciton function
+% doComputeLoss     Set to true to compute loss
 
 trRMSE = zeros(1, groups);
 teRMSE = zeros(1, groups);
 
-if computeLoss
+if doComputeLoss
     trLoss = zeros(1, groups);
     teLoss = zeros(1, groups);
     trLogLoss = zeros(1, groups);
@@ -38,7 +38,7 @@ for groupNumber = 1:groups
     teRMSE(groupNumber) = computeRMSE(yTe, yTePred);
     
     % Loss function
-    if computeLoss
+    if doComputeLoss
         trLoss(groupNumber) = computeLoss(yTr, yTrPred);
         teLoss(groupNumber) = computeLoss(yTe, yTePred);
         
@@ -50,7 +50,7 @@ end
 trError.RMSE = trRMSE;
 teError.RMSE = teRMSE;
 
-if computeLoss
+if doComputeLoss
     trError.Loss = trLoss;
     teError.Loss = teLoss;
     

### 2015-11-02 report/paper.tex

@@
 	\center
-	\includegraphics[width=2.6in]{figures/y_histogram.png}
+	\includegraphics[width=2.2in]{figures/y_histogram.png}
 	\caption{Histogram of $y$. We can see three bursts of data.}
@@
 \center
-\subfigure[$y$ in function of $55^{th}$ dimension of $X$.]{\includegraphics[width=2.6in]{figures/dim6.png} \label{fig:reg_dim6}}
+\subfigure[$y$ in function of $55^{th}$ dimension of $X$.]{\includegraphics[width=2.2in]{figures/dim6.png} \label{fig:reg_dim6}}
 \hfill
-\subfigure[$y$ in function of $6^{th}$ dimension of $X$.]{\includegraphics[width=2.6in]{figures/dim55.png} \label{fig:reg_dim55}}
+\subfigure[$y$ in function of $6^{th}$ dimension of $X$.]{\includegraphics[width=2.2in]{figures/dim55.png} \label{fig:reg_dim55}}
 \caption{Graphs of $y$ in function of the input dimensions most correlated with the output.}
@@
 	\center
-	\includegraphics[width=2.6in]{figures/clusters_train.png}
+	\includegraphics[width=2.2in]{figures/clusters_train.png}
 	\caption{Clusters obtained along the two most correlated dimensions. We can see three clearly delimited groups of data.}
@@
 	\center
-	\subfigure[First k-means applied along the most correlated dimension of $X$ with $y$.]{\includegraphics[width=2.6in]{figures/clusters_phase1.png} \label{fig:reg_clustersTest1}}
+	\subfigure[First k-means applied along the most correlated dimension of $X$ with $y$.]{\includegraphics[width=2.2in]{figures/clusters_phase1.png} \label{fig:reg_clustersTest1}}
 	\hfill
-	\subfigure[Second k-means applied along the second most correlated dimension of $X$ with $y$.]{\includegraphics[width=2.6in]{figures/clusters_phase2.png} \label{fig:reg_clustersTest2}}
+	\subfigure[Second k-means applied along the second most correlated dimension of $X$ with $y$.]{\includegraphics[width=2.2in]{figures/clusters_phase2.png} \label{fig:reg_clustersTest2}}
 	\caption{Steps of k-means clustering for test-data.}
@@
 Cross-validation was made using $66 \% $ of the data as train data and the rest as test data.
+We observe that the bias is higher and the variance globally lower for small values of $\alpha$.
+The RMSE is optimal for $\alpha \approx 0.2$.
+With higher $\alpha$ values, gradient descent is not able to converge and the error goes to $\infty$.
 
@@
 	\center
-	\includegraphics[width=4in]{figures/lambdaRMSE.png}
+	\includegraphics[width=3in]{figures/alphaRMSEBox.png}
 	\caption{Ridge regression for a $66 \%$ - $33 \%$ split with 3 iterations for each $\lambda$.}
@@
 We can see that $\lambda$ should be chosen between $0$ and $100$ to have the best prediction and that $\lambda$ higher than $1000$ increases the RMSE and the model under-fits.
+For high $\lambda$ values, the model and estimation bias are increased as their variance is reduced.
+
+\begin{figure}[!h] % !t
+	\center
+	\includegraphics[width=3in]{figures/lambdaRMSE.png}
+	\caption{Ridge regression for a $66 \%$ - $33 \%$ split with 3 iterations for each $\lambda$.}
+	\label{fig:reg_lambdaRMSE}
+\end{figure}
+
+Figure ~\ref{fig:reg_compareMethods} compares three regression methods with $\alpha = 0.25$ for the gradient descent and $\lambda = 100$ for ridge regression. We see that TODO.
 
@@
 	\includegraphics[width=4in]{figures/lambdaRMSE.png}
-	\caption{Ridge regression for a $66 \%$ - $33 \%$ split with 3 iterations for each $\lambda$.}
-	\label{fig:reg_lambdaRMSE}
+	\caption{Boxplots of train and test RMSE for least-squares, least-square with gradient descent and ridge regression.}
+	\label{fig:reg_compareMethods}
 \end{figure}
-
-Figure TODO compares the three regression methods with $\alpha = TODO$ for the gradient descent and $\lambda = TODO$ for ridge regression. We see that TODO.
 
@@
 \section{Summary}
-In this report, we ...
+In this report, we analyse a regression and a classification dataset. We used k-means clustering and PCA to preprocess the regression data. We found that TODO.
+We separated the classification data using a binary dimension that classify the data in two different distributions along the variable of the input data.We found that TODO.
 

### 2015-11-02 report/paper.tex
Update report files

@@
 As the input matrix is ill-conditioned, we applied a PCA algorithm to reduce its dimensionality.
-WHICH ALGORITHM IS THE BEST; IDEA OF ERROR.
+We observe that the three methods are equivalent, with RMSE around $520$ for train and for test parts.
 For the classification part, we separate the data given their distribution along the dimensions of the input.
@@
 
-Figure ~\ref{fig:reg_compareMethods} compares three regression methods with $\alpha = 0.25$ for the gradient descent and $\lambda = 100$ for ridge regression. We see that TODO.
+Figure ~\ref{fig:reg_compareMethods} compares three regression methods with $\alpha = 0.25$ for the gradient descent and $\lambda = 100$ for ridge regression.
+We see that the three methods give equivalent results.
+This is surely the consequence of preprocessing the data with PCA, so that the input matrix is not ill conditioned.
 
@@
 	\center
-	\includegraphics[width=4in]{figures/lambdaRMSE.png}
+	\includegraphics[width=4in]{figures/comparisonReg.png}
 	\caption{Boxplots of train and test RMSE for least-squares, least-square with gradient descent and ridge regression.}
@@
 In this report, we analyse a regression and a classification dataset. We used k-means clustering and PCA to preprocess the regression data. We found that TODO.
-We separated the classification data using a binary dimension that classify the data in two different distributions along the variable of the input data.We found that TODO.
+We separated the classification data using a binary dimension that classify the data in two different distributions along the variable of the input data. We found that TODO.
 

### 2015-11-02 src/regression/generatePlots.m
Update plots

@@ -229,3 +229,15 @@ for i = 1:numIterations
     eastSquaresGDTeRMSE(i) = mean(teError.RMSE);
 end
 
+%%
+figure()
+ax(1) = subplot(121);
+boxplot([leastSquaresTrRMSE' leastSquaresGDTrRMSE' ridgeTrRMSE'],'labels',{'least-squares','least-squares GD','ridge regression'});
+ylabel('Train RMSE')
+ax(2) = subplot(122);
+boxplot([leastSquaresTeRMSE' eastSquaresGDTeRMSE' ridgeTeRMSE'],'labels',{'least-squares','least-squares GD','ridge regression'});
+ylabel('Test RMSE')
+
+
+
+

### 2015-11-02 src/regression/main.m
Generate results

@@ -7,16 +7,25 @@ addpath('../methods');
 
 load('../../data/SaoPaulo_regression.mat')
 
-% predict = @(Xtr, Ytr, Xte) predictRegression(Xtr, Ytr, Xte, 10^5);
-% [ trRMSE, teRMSE ] = crossValidation(y_train, X_train, 3, predict);
 
-[ trRMSE, teRMSE, lambda ] = findBestRidgeLambda(X_train, y_train);
-%%
-figure()
-ax(1) = subplot(211);
-boxTr = boxplot(trRMSE,'plotstyle','compact','colors','b','labels',lambda);
-ax(2) = subplot(212);
-hold on
-boxTe = boxplot(teRMSE,'plotstyle','compact','colors','r');
-linkaxes(ax);
-%ylim([0.05 0.12])
+predictor = @(y, tX) ridgeRegression(y, tX, 100);
+predict = @(Xtr, ytr, Xte) predictRegression(Xtr, ytr, Xte, predictor);
+
+yTe = predict(X_train, y_train, X_test);
+csvwrite('results/predictions_regression.csv', yTe);
+
+
+nbIteration = 100;
+
+RMSE = zeros(1, nbIteration);
+for i = 1:nbIteration
+    home;
+    fprintf('Iteration %d out of %d\n', i, nbIteration);
+    [~, teError] = crossValidation(X_train, y_train, 3, predict, false);
+    RMSE(i) = mean(teError.RMSE);
+end
+
+file = fopen('results/test_errors_regression.csv', 'w');
+fprintf(file, 'rmse,%d', round(mean(RMSE)));
+fclose(file);
+

### 2015-11-02 src/regression/generatePlots.m
Update report and plots

@@ -194,7 +194,7 @@ close;
 %% Plot ridge vs leastSquares vs leastSquaresGD
 
 lambda = 100;
-alpha = 0.25;
+alpha = 0.3;
 groups = 3;
 numIterations = 100;
 
@@ -206,11 +206,11 @@ ridgePredict = @(Xtr, ytr, Xte){
     predictRegression(Xtr, ytr, Xte, ridgePredictor);
 };
 leastSquaresPredict = @(Xtr, ytr, Xte){
-    predictRegression(Xtr, ytr, Xte, ridgePredictor);
+    predictRegression(Xtr, ytr, Xte, leastSquaresPredictor);
 };
 
 leastSquaresGDPredict = @(Xtr, ytr, Xte){
-    predictRegression(Xtr, ytr, Xte, ridgePredictor);
+    predictRegression(Xtr, ytr, Xte, leastSquaresGDPredictor);
 };
 
 
@@ -226,18 +226,21 @@ for i = 1:numIterations
     
     [trError, teError] = crossValidation(X_train, y_train, groups, leastSquaresGDPredict, false);
     leastSquaresGDTrRMSE(i) = mean(trError.RMSE);
-    eastSquaresGDTeRMSE(i) = mean(teError.RMSE);
+    leastSquaresGDTeRMSE(i) = mean(teError.RMSE);
 end
 
 %%
 figure()
 ax(1) = subplot(121);
-boxplot([leastSquaresTrRMSE' leastSquaresGDTrRMSE' ridgeTrRMSE'],'labels',{'least-squares','least-squares GD','ridge regression'});
+boxplot([leastSquaresTrRMSE leastSquaresGDTrRMSE' ridgeTrRMSE'],...
+    'labels',{'LS','LS-GD','Ridge'});
 ylabel('Train RMSE')
+ylim([400 700])
 ax(2) = subplot(122);
-boxplot([leastSquaresTeRMSE' eastSquaresGDTeRMSE' ridgeTeRMSE'],'labels',{'least-squares','least-squares GD','ridge regression'});
+boxplot([leastSquaresTeRMSE leastSquaresGDTeRMSE' ridgeTeRMSE'],...
+    'labels',{'LS','LS-GD','Ridge'});
 ylabel('Test RMSE')
-
+ylim([400 700])
 
 
 

### 2015-11-02 report/paper.tex

@@
 As the input matrix is ill-conditioned, we applied a PCA algorithm to reduce its dimensionality.
-We observe that the three methods are equivalent, with RMSE around $520$ for train and for test parts.
+We observe that TODO.
+We obtained similar train and test RMSE around $520$.
 For the classification part, we separate the data given their distribution along the dimensions of the input.
@@
 
-Figure ~\ref{fig:reg_compareMethods} compares three regression methods with $\alpha = 0.25$ for the gradient descent and $\lambda = 100$ for ridge regression.
-We see that the three methods give equivalent results.
-This is surely the consequence of preprocessing the data with PCA, so that the input matrix is not ill conditioned.
+Figure ~\ref{fig:reg_compareMethods} compares three regression methods with $\alpha = 0.3$ for the gradient descent and $\lambda = 100$ for ridge regression.
+It is important to note that some outliers (11 for the train, 17 for the test) of least-squares method are not visible in this graph, as their RMSE is higher than $10^4$.
 
@@
 	\center
-	\includegraphics[width=4in]{figures/comparisonReg.png}
-	\caption{Boxplots of train and test RMSE for least-squares, least-square with gradient descent and ridge regression.}
+	\includegraphics[width=3in]{figures/comparisonReg.png}
+	\caption{Boxplots of train and test RMSE for least-squares (LS), least-squares with gradient descent (LS-GD) and ridge regression (Ridge).}
 	\label{fig:reg_compareMethods}
 \end{figure}
+
+The median train/test RMSE are respectively equal to $526/539$, $551/562$ and $526/524$ for least-squares, least-squares GD and ridge regression.
+We observe that least-squares GD has a high bias and a low variance compared to least-squares that has a lower bias and a higher variance.
 
@@
 \section{Summary}
-In this report, we analyse a regression and a classification dataset. We used k-means clustering and PCA to preprocess the regression data. We found that TODO.
+In this report, we analyse a regression and a classification dataset.
+We used k-means clustering and PCA to preprocess the regression data.
+We found that least-squares, least-squares with gradient descent and ridge regression TO DO.
+
 We separated the classification data using a binary dimension that classify the data in two different distributions along the variable of the input data. We found that TODO.

### 2015-11-02 src/classification/findBestPenLambda.m
CHICHI

@@ -2,7 +2,7 @@ function [ trRMSE, teRMSE, trLoss, teLoss, lambdas ] = findBestPenLambda( X, y )
 %FINDBESTRIDGELAMBDA Summary of this function goes here
 %   Detailed explanation goes here
 
-lambdas = logspace(-3, 7, 60);
+lambdas = logspace(-3, 0, 10);
 
 predictParam = @(XTr, yTr, XTe, param) {
     predictClassification(XTr, yTr, XTe, @(XTr, yTr) penLogisticRegression(XTr, yTr, 0.001, param));

### 2015-11-02 src/classification/generatePlots.m

@@ -19,6 +19,16 @@ setFontSize = @(gca, gcf, xl, yl) {
     set(xl, 'FontSize', LABEL_FONT_SIZE); set(yl, 'FontSize', LABEL_FONT_SIZE)
     };
 
+setFontSizeNoTouchTicks = @(gca, gcf, xl, yl) {
+    %set(gca, 'FontSize', FONT_SIZE); 
+    %set(findall(gcf,'type','text'), 'FontSize', FONT_SIZE); 
+    set(xl, 'FontSize', LABEL_FONT_SIZE); set(yl, 'FontSize', LABEL_FONT_SIZE);
+    };
+
+getShortLabels = @(nums) {
+    arrayfun(@(x) sprintf('%.1E', x), nums, 'UniformOutput', false);
+    };
+
 
 %% Histogram of dimension 35
 figure();
@@ -64,5 +74,24 @@ yl = ylabel('Frequency');
 setFontSize(gca, gcf, xl, yl);
 
 print -dpng 'plots/dim23.png';
+close;
+
+%% BoxPlots of lambda vs Loss
+
+[ trRMSE, teRMSE, trLoss, teLoss, lambda ] = findBestPenLambda(X_train, y_train);
+
+%%
+figure()
+boxplot(trLoss,'plotstyle','compact','colors','b','labels', getShortLabels(lambda));
+hold on
+boxplot(teLoss,'plotstyle','compact','colors','r', 'labels', getShortLabels(lambda));
+xl = xlabel('Lambda');
+yl = ylabel('Loss');
+
+setFontSizeNoTouchTicks(gca, gcf, xl, yl);
+
+ylim([0 0.15])
+
+print -dpng 'plots/lambdaLossBox.png';
 close;
 

### 2015-11-02 src/classification/main.m

@@ -10,17 +10,3 @@ load('../../data/SaoPaulo_classification.mat')
 % predict = @(Xtr, Ytr, Xte) predictRegression(Xtr, Ytr, Xte, 10^5);
 % [ trRMSE, teRMSE ] = crossValidation(y_train, X_train, 3, predict);
 
-[ trRMSE, teRMSE, trLoss, teLoss, lambda ] = findBestPenLambda(X_train, y_train);
-
-% semilogx(lambda, teRMSE, 'r*-')
-% hold on
-% semilogx(lambda, trRMSE, 'b*-')
-
-%%
-figure()
-% subplot(211)
-boxTr = boxplot(trRMSE,'plotstyle','compact','colors','b','labels',lambda);
-% subplot(212)
-hold on
-boxTe = boxplot(teRMSE,'plotstyle','compact','colors','r');
-ylim([0.05 0.12])

### 2015-11-02 report/paper.tex

@@
 As the input matrix is ill-conditioned, we applied a PCA algorithm to reduce its dimensionality.
-We observe that TODO.
-We obtained similar train and test RMSE around $520$.
+Ridge regression is preferred over least-squares and least-squares with gradient descent, as it has the lower bias and variance.
 For the classification part, we separate the data given their distribution along the dimensions of the input.
-WHICH ALGORITHM IS THE BEST; IDEA OF ERROR.
+WHICH ALGORITHM IS THE BEST.
 \end{abstract}
@@
 We observe that least-squares GD has a high bias and a low variance compared to least-squares that has a lower bias and a higher variance.
+Even if least-squares and ridge regression have similar train and test RMSE, we prefer to use ridge regression as its variance is lower and as its outliers are not very far from the median compared to those of least-squares.
 
@@
 
-Figure TODO shows the train and test TODO ERROR for $\lambda$ varied from
+Figure ~\ref{fig:class_penLogistic} shows the train and test loss function for $\lambda$ varied from $10^{-3}$ to $1$.
+
+\begin{figure}[!h] % !t
+	\center
+	\includegraphics[width=3in]{figures/comparisonReg.png}
+	\caption{Boxplots of train and test loss function for penalised logistic regression in function of $\lambda$.}
+	\label{fig:class_penLogistic}
+\end{figure}
 
@@
 We used k-means clustering and PCA to preprocess the regression data.
-We found that least-squares, least-squares with gradient descent and ridge regression TO DO.
+We found that ridge regression give a reasonable fit with a test RMSE equal to $524 \pm 20$.
+Least-squares and least-squares with gradient descent predictions have respectively a higher variance and a higher bias compared to ridge regression.
 
@@
 \subsubsection*{Acknowledgments}
-...
+We would like to thank Zo\'e Baraschi for her support.
+She was there to help us in moments of darkness with food and hugs.
+We also would like to thank moodle to have given us this half an hour that was unhoped for.
+It gave us the courage and time to finish writing up this report.
+
+We also collaborated efficiently to generate all the code and the report.
 

### 2015-11-02 src/classification/main.m
CHICHI giving up for tonight

@@ -7,6 +7,23 @@ addpath('../methods');
 
 load('../../data/SaoPaulo_classification.mat')
 
-% predict = @(Xtr, Ytr, Xte) predictRegression(Xtr, Ytr, Xte, 10^5);
-% [ trRMSE, teRMSE ] = crossValidation(y_train, X_train, 3, predict);
+predictor = @(y, tX) penLogisticRegression(y, tX, 0.01);
+predict = @(Xtr, ytr, Xte) predictClassification(Xtr, ytr, Xte, predictor);
 
+[yTe, pTe] = predict(X_train, y_train, X_test);
+csvwrite('results/predictions_classification.csv', pTe);
+
+
+nbIteration = 3;
+
+RMSE = zeros(1, nbIteration);
+for i = 1:nbIteration
+    home;
+    fprintf('Iteration %d out of %d\n', i, nbIteration);
+    [~, teError] = crossValidation(X_train, y_train, 3, predict, false);
+    RMSE(i) = mean(teError.RMSE);
+end
+
+file = fopen('results/test_errors_regression.csv', 'w');
+fprintf(file, 'loss,%d', round(mean(RMSE)));
+fclose(file);

### 2015-11-02 report/paper.tex
Update report

@@
 \center
-\subfigure[$y$ in function of $55^{th}$ dimension of $X$.]{\includegraphics[width=2.2in]{figures/dim6.png} \label{fig:reg_dim6}}
+\subfigure[$y$ in function of $55^{th}$ dimension of $X$.]{\includegraphics[width=2in]{figures/dim6.png} \label{fig:reg_dim6}}
 \hfill
-\subfigure[$y$ in function of $6^{th}$ dimension of $X$.]{\includegraphics[width=2.2in]{figures/dim55.png} \label{fig:reg_dim55}}
+\subfigure[$y$ in function of $6^{th}$ dimension of $X$.]{\includegraphics[width=2in]{figures/dim55.png} \label{fig:reg_dim55}}
 \caption{Graphs of $y$ in function of the input dimensions most correlated with the output.}
@@
 Cross-validation was made using $66 \% $ of the data as train data and the rest as test data.
-We observe that the bias is higher and the variance globally lower for small values of $\alpha$.
-The RMSE is optimal for $\alpha \approx 0.2$.
+Train and test RMSE is optimal for $\alpha \approx 0.2$.
 With higher $\alpha$ values, gradient descent is not able to converge and the error goes to $\infty$.
@@
 	\center
-	\includegraphics[width=3in]{figures/alphaRMSEBox.png}
-	\caption{Ridge regression for a $66 \%$ - $33 \%$ split with 3 iterations for each $\lambda$.}
+	\includegraphics[width=2.6in]{figures/alphaRMSE.png}
+	\caption{Least-squares with gradient descent for a $66 \%$ - $33 \%$ split with 3 iterations for each $\alpha$.}
 	\label{fig:reg_alphaRMSE}
@@
 	\center
-	\includegraphics[width=3in]{figures/lambdaRMSE.png}
+	\includegraphics[width=2.6in]{figures/lambdaRMSE.png}
 	\caption{Ridge regression for a $66 \%$ - $33 \%$ split with 3 iterations for each $\lambda$.}
@@
 	\includegraphics[width=3in]{figures/comparisonReg.png}
-	\caption{Boxplots of train and test RMSE for least-squares (LS), least-squares with gradient descent (LS-GD) and ridge regression (Ridge).}
+	\caption{Boxplots of train and test RMSE for least-squares (LS), least-squares with gradient descent (LS-GD) and ridge regression (Ridge) with $100$ iterations for each method.}
 	\label{fig:reg_compareMethods}
@@
 	\center
-	\subfigure[Histogram of the $23^{th}$ dimension of $X$.]{\includegraphics[width=2.6in]{figures/dim23.png} \label{fig:class_dim23}}
+	\subfigure[Histogram of the $23^{th}$ dimension of $X$.]{\includegraphics[width=2.1in]{figures/dim23.png} \label{fig:class_dim23}}
 	\hfill
-	\subfigure[Histogram of the $35^{th}$ dimension of $X$.]{\includegraphics[width=2.6in]{figures/dim35.png} \label{fig:class_dim35}}
+	\subfigure[Histogram of the $35^{th}$ dimension of $X$.]{\includegraphics[width=2.1in]{figures/dim35.png} \label{fig:class_dim35}}
 	\caption{Example of histogram to show the distribution of $X$ dimensions.
@@
 Figure ~\ref{fig:class_penLogistic} shows the train and test loss function for $\lambda$ varied from $10^{-3}$ to $1$.
+We observe that changing $\lambda$ does not really improve the train and test loss function.
+The difference of error between train and test shows that the model is over-fitting.
+This is surely due to the amount of train data compared to the amount of test data, which is not high enough.
+However, we chose this split to make a cross-validation with similar train/test quantities to those we are provided to make output prediction.
 
@@
 	\center
-	\includegraphics[width=3in]{figures/comparisonReg.png}
-	\caption{Boxplots of train and test loss function for penalised logistic regression in function of $\lambda$.}
+	\includegraphics[width=2.6in]{figures/lambdaLossBox.png}
+	\caption{Penalised logistic regression for a $50 \% - 50\%$ split with 3 iterations for each $\lambda$.}
 	\label{fig:class_penLogistic}
@@
 \subsubsection*{References}
+Applied Machine Learning course notes from Aude Billard
+
+PCML course notes from Mohammad Emtiyaz Khan
 

### 2015-11-03 src/classification/findBestPenLambda.m
Update report, logistic regression, code for submission

@@ -1,6 +1,6 @@
 function [ trRMSE, teRMSE, trLoss, teLoss, lambdas ] = findBestPenLambda( X, y )
-%FINDBESTRIDGELAMBDA Summary of this function goes here
-%   Detailed explanation goes here
+%FINDBESTRIDGELAMBDA Find the optimal lambda to minimise the error of the
+%penalised logistic regression
 
 lambdas = logspace(-3, 0, 10);
 

### 2015-11-03 src/classification/generatePlots.m

@@ -98,8 +98,8 @@ close;
 %% BoxPlots of PenLogistic Regression vs Logistic
 
 lambda = 0.01;
-groups = 3;
-numIterations = 2;
+groups = 2;
+numIterations = 50;
 
 penLogisticPredictor = @(y, tX) penLogisticRegression(y, tX, 0.001, lambda);
 logisticPredictor = @(y, tX) logisticRegression(y, tX, 0.001);
@@ -124,30 +124,22 @@ for i = 1:numIterations
     [trError, teError] = crossValidation(X_train, y_train, groups, penLogPredict, true);
     penLogTrLoss(i) = mean(trError.Loss);
     penLogTeLoss(i) = mean(teError.Loss);
-    disp(size(trError))
-    trError
-    disp(size(penLogTrLoss))
-    penLogTrLoss
     
     [trError, teError] = crossValidation(X_train, y_train, groups, logPredict, true);
     logTrLoss(i) = mean(trError.Loss);
     logTeLoss(i) = mean(teError.Loss);
-    disp(size(trError))
-    trError
-    disp(size(logTrLoss))
-    logTrLoss
     fprintf('End of iteration %d\n', i);
 end
 
 %%
 figure()
 ax(1) = subplot(121);
-boxplot([penLogTrLoss' logTrLoss'],...
-    'labels',{'PenLog','Log'});
-ylabel('Train RMSE')
-%ylim([400 700])
+boxplot([logTrLoss' penLogTrLoss'],...
+    'labels',{'Log','PenLog'});
+ylabel('Train Loss Function')
+ylim([0.03 0.16])
 ax(2) = subplot(122);
-boxplot([penLogTeLoss' logTeLoss'],...
-    'labels',{'PenLog','Log'});
-ylabel('Test RMSE')
-%ylim([400 700])
+boxplot([logTeLoss' penLogTeLoss'],...
+    'labels',{'Log','PenLog'});
+ylabel('Test Loss Function')
+ylim([0.03 0.16])

### 2015-11-03 src/classification/main.m

@@ -7,23 +7,23 @@ addpath('../methods');
 
 load('../../data/SaoPaulo_classification.mat')
 
-predictor = @(y, tX) penLogisticRegression(y, tX, 0.01);
+predictor = @(y, tX) penLogisticRegression(y, tX, 0.001, 0.01);
 predict = @(Xtr, ytr, Xte) predictClassification(Xtr, ytr, Xte, predictor);
 
 [yTe, pTe] = predict(X_train, y_train, X_test);
 csvwrite('results/predictions_classification.csv', pTe);
 
 
-nbIteration = 3;
+nbIteration = 50;
 
-RMSE = zeros(1, nbIteration);
+lossFunction = zeros(1, nbIteration);
 for i = 1:nbIteration
     home;
     fprintf('Iteration %d out of %d\n', i, nbIteration);
-    [~, teError] = crossValidation(X_train, y_train, 3, predict, false);
-    RMSE(i) = mean(teError.RMSE);
+    [~, teError] = crossValidation(X_train, y_train, 2, predict, true);
+    lossFunction(i) = mean(teError.Loss);
 end
 
-file = fopen('results/test_errors_regression.csv', 'w');
-fprintf(file, 'loss,%d', round(mean(RMSE)));
+file = fopen('results/test_errors_classification.csv', 'w');
+fprintf(file, 'loss,%1.3f', mean(lossFunction));
 fclose(file);

### 2015-11-03 src/classification/predictClassification.m

@@ -44,8 +44,6 @@ for cl = 1:sizeCluster
     yTrF{cl,1} = yTrFiltered;
     
     % Compute (penalised) logistic regression
-%     beta{cl,1} = logisticRegression(yTrF{cl,1}, tXTr{cl,1}, 0.001);
-%    beta{cl,1} = penLogisticRegression(yTrF{cl,1}, tXTr{cl,1}, 0.001, 0.01);
     beta{cl,1} = localPredictor(yTrF{cl,1}, tXTr{cl,1});
     
 end

### 2015-11-03 src/methods/leastSquaresGD.m

@@ -1,7 +1,7 @@
 function beta = leastSquaresGD( y, tX, alpha )
 
 % Maximum number of iterations
-maxIters = 10000; % TO DEFINE
+maxIters = 10000;
 epsilon = alpha * 1.5;
 
 % Number of data
@@ -23,8 +23,8 @@ for k = 1:maxIters
     beta = beta - alpha.*g;
     
     % Store beta and L 
-    beta_all(:,k) = beta; % FOR DEBUG
-    L_all(k) = L; % FOR DEBUG
+    beta_all(:,k) = beta; 
+    L_all(k) = L; 
     
     % Look at the convergence
     if k>1

### 2015-11-03 src/methods/logisticRegression.m

@@ -1,7 +1,7 @@
 function beta = logisticRegression( y, tX, alphaInitial )
 
 % Maximum number of iterations
-maxIters = 15000; % TO DEFINE
+maxIters = 15000;
 
 % Number of data
 N = size(tX,1);
@@ -19,24 +19,28 @@ for k = 1:maxIters
     % Compute the Hessian
     H = computeHessian(tX, beta);
     
-    % Update beta using gradient descent
-    % beta = beta - alpha.*g;
     % Update beta using Newton's method
     beta = beta - alpha.*(H^-1)*g;
     
     % Compute cost of logistic regression
     L = computeCostLogisticReg(y, tX, beta);
+    
+    % Break if the algorithm has diverged during the current iteration
+    if L == Inf
+        fprintf('Cost explosion, last iteration taken\n');
+        beta_all(:,k) = beta_all(:,k-1);
+        break;
+    end
+    
+    % Compute step size
     alpha = abs(L*10^(-4));
     
     % Store beta and L
     beta_all(:,k) = beta;
     L_all(k) = L;
     
-    
-    
     % Look at the convergence
     if k>1
-%         fprintf('%d : %f  %f %f\n', k,L, alpha, max(abs(beta-beta_all(:,k-1))));
         if abs(beta-beta_all(:,k-1))<0.001
             % If the difference between last and present beta is small...
             % break

### 2015-11-03 src/methods/penLogisticRegression.m

@@ -1,7 +1,7 @@
 function beta = penLogisticRegression(y,tX,alphaInitial,lambda)
 
 % Maximum number of iterations
-maxIters = 5000; % TO DEFINE
+maxIters = 5000; 
 
 % Number of data
 N = size(tX,1);
@@ -20,15 +20,14 @@ for k = 1:maxIters
     % Compute the Hessian
     H = computePenHessian(tX, beta, lambda);
     
-    % Update beta using gradient descent
-    % beta = beta - alpha.*g;
     % Update beta using Newton's method
     beta = beta - alpha.*(H^-1)*g;
     
     % Compute cost of logistic regression
     L = computeCostPenLogisticReg(y, tX, beta, lambda);
-    alpha = abs(L*10^(-4));
     
+    % Compute step size
+    alpha = abs(L*10^(-4));
     
     % Store beta and L
     beta_all(:,k) = beta;
@@ -36,7 +35,6 @@ for k = 1:maxIters
     
     % Look at the convergence
     if k>1
-%         fprintf('%d : %f  %f %f\n', k,L, alpha, max(abs(beta-beta_all(:,k-1))));
         if abs(beta-beta_all(:,k-1))<0.001
             % If the difference between last and present beta is small...
             % break

### 2015-11-03 src/regression/findBestGDAlpha.m

@@ -1,6 +1,5 @@
 function [ trRMSE, teRMSE, lambdas ] = findBestGDAlpha( X, y )
-%FINDBESTRIDGELAMBDA Summary of this function goes here
-%   Detailed explanation goes here
+%FINDBESTGDALPHA 
 
 lambdas = logspace(-3, -0.5, 20);
 

### 2015-11-03 src/regression/findBestRidgeLambda.m

@@ -1,6 +1,5 @@
 function [ trRMSE, teRMSE, lambdas ] = findBestRidgeLambda( X, y )
-%FINDBESTRIDGELAMBDA Summary of this function goes here
-%   Detailed explanation goes here
+%FINDBESTRIDGELAMBDA
 
 lambdas = logspace(-3, 7, 60);
 

### 2015-11-03 src/regression/predictRegression.m

@@ -18,12 +18,6 @@ numClusters = 3;
 
 clustersPred = predictClusters(XTe);
 
-%figure()
-%plot(XTe(clustersPred == 1, clusteredDim(1)),XTe(clustersPred == 1, clusteredDim(2)),'r*');
-%hold on
-%plot(XTe(clustersPred == 2, clusteredDim(1)),XTe(clustersPred == 2, clusteredDim(2)),'b*');
-%plot(XTe(clustersPred == 3, clusteredDim(1)),XTe(clustersPred == 3, clusteredDim(2)),'g*');
-
 %% Train regression model for each cluster
 V = cell(numClusters, 1);
 VReduced = cell(numClusters, 1);
@@ -50,16 +44,9 @@ for cl = 1:numClusters
     % Compute tX and final y
     tXTr{cl,1} = [ones(length(XTrKept), 1)  XTrKept];
     yTrF{cl,1} = yTrFiltered;
-    
-    % Compute regression
-%     beta{cl,1} = leastSquaresGD(yTrF{cl,1}, tXTr{cl,1}, param);
-%     beta{cl,1} = leastSquares(yTrF{cl,1}, tXTr{cl,1});
- %   beta{cl,1} = ridgeRegression(yTrF{cl,1}, tXTr{cl,1}, param);
  
     beta{cl, 1} = localPredictor(yTrF{cl,1}, tXTr{cl,1});
  
-%     figure()
-%     plot(XTrKept(:,1),yTrF{cl,1},'*');
 end
 
 %% Test 

### 2015-11-05 report/paper.tex
Update last changes for the report

@@
 \begin{abstract}
-In this report, we study several linear method to compute regression and classification on two different datasets.
+In this report, we study several linear methods to compute regression and classification on two different datasets.
 For the regression part, we used k-means clustering to separate the data in three clusters.
@@
 Ridge regression is preferred over least-squares and least-squares with gradient descent, as it has the lower bias and variance.
-For the classification part, we separate the data given their distribution along the dimensions of the input.
-WHICH ALGORITHM IS THE BEST.
+For the classification part, we separated the data given their distribution along the dimensions of the input.
+Penalised logistic regression is preferred over logistic regression as it converges more reliably to a model and has lower bias and variance.
 \end{abstract}
@@
 Figure ~\ref{fig:reg_yHistogram} shows the histogram of the output variable. We
-see that the data are split in three bursts of different sizes : a large amount
+see that the data is split in three bursts of different sizes : a large amount
 of the data have small output values and only a small part have the highest
@@
 
-\begin{figure}[!h] % !t
-	\center
-	\includegraphics[width=2.2in]{figures/y_histogram.png}
-	\caption{Histogram of $y$. We can see three bursts of data.}
-	\label{fig:reg_yHistogram}
-\end{figure}
-
 We also computed the eigenvalues of the matrix $X^T X$.
@@
 \center
-\subfigure[$y$ in function of $55^{th}$ dimension of $X$.]{\includegraphics[width=2in]{figures/dim6.png} \label{fig:reg_dim6}}
+\subfigure[Histogram of $y$. We can see three bursts of data.]{\includegraphics[width=1.8in]{figures/y_histogram.png} \label{fig:reg_yHistogram}}
 \hfill
-\subfigure[$y$ in function of $6^{th}$ dimension of $X$.]{\includegraphics[width=2in]{figures/dim55.png} \label{fig:reg_dim55}}
-\caption{Graphs of $y$ in function of the input dimensions most correlated with the output.}
+\subfigure[$y$ in function of the most correlated dimension ($55^{th}$) of $X$ with $y$.]{\includegraphics[width=1.8in]{figures/dim55.png} \label{fig:reg_dim6}}
+\hfill
+\subfigure[$y$ in function of the second most correlated dimension ($6^{th}$) of $X$ with $y$.]{\includegraphics[width=1.8in]{figures/dim6.png} \label{fig:reg_dim55}}
+\caption{Regression data visualisation.}
 \end{figure}
@@
 During the initialisation, we sort the data-points by their norm and separate the sorted list in $k$ chunks, where $k$ is the given number of clusters.
-The mean along each dimension is then calculated for each chunk giving the initial centres for the clusters.
+The mean along each dimension is then calculated for each chunk and gives the position of each initial cluster centre.
 
-During the test part, we follow the same principle.
-The most correlated dimension of $X$ found during the training part is used to find the closest centre and the second most correlated dimension is then used to divide the data classified in the sparse cluster in two sub-clusters.
-
-Figure ~\ref{fig:reg_clustersTest} shows an example of k-means clustering for test data.
+During the test part, we follow the same principle as for the training.
+First, the most correlated dimension of $X$ found during the training part is used to assign each point to one of the two big clusters, as shown in Figure ~\ref{fig:reg_clustersTest1}.
+The second most correlated dimension is then used to divide the data classified in the sparse cluster in two sub-clusters, as shown in Figure ~\ref{fig:reg_clustersTest2}.
 
@@
 We defined as outliers the data-points for which some elements of one or more real-valued dimensions are further than 3 standard deviations from the mean.
-As our data were normalised, all non-categorical element higher than 3 were considered as outlier elements.
+As our data were normalised, all non-categorical elements higher than 3 were considered as outlier elements.
 To fix the outliers, value $3$ (or $-3$) was assigned to those elements.
-Fixing the outliers with this methods allows us to avoid errors in prediction models due to false-classified data.
+Fixing the outliers with this method allows us to limit errors due to false-classified data in our model.
 
@@
 It projects our input data in a new coordinate system based on the covariance between dimensions.
-It allows us to reduce the dimensionality of our data in this new space by removing the dimensions with eigenvalues lower than a defined value.
-We choose this value equal to $1$ as eleven dimensions have eigenvalues in the order of $10^{-10}$ and all others are higher than $10$.
+It allows us to reduce the dimensionality of our data in this new space by removing the dimensions with eigenvalues lower than a defined threshold.
+We choose this threshold equal to $1$ as eleven dimensions have eigenvalues in the order of $10^{-10}$ and all others are higher than $10$.
 The data are then projected back in the initial coordinate system.
-In our case, the PCA reduce our input data by eleven dimensions.
+In our case, the PCA reduces our input data by eleven dimensions.
 
 After the clustering of test-data, each group is also treated separately.
-The data are first normalised given the means and standard deviations computed during the training.
+The data is first normalised given means and standard deviations computed during the training.
 The outliers are then fixed and PCA is applied by projecting the data with the eigenvectors computed during the training part.
@@
 Using PCA allows us to solve the problem of ill-conditioned matrix, so that we can use the three methods mentioned above.
-The ridge regression is the only method that can be used without reducing the dimensionality of the input data.
+Ridge regression is the only method that can be used without reducing the dimensionality of our input data.
 
-We computed the RMSE for each model with and without the categorical dimensions of X.
-As removing those dimensions does not change significantly the error, we decided to compute the regression only with the 60 real-valued dimensions.
+We computed RMSE for each model with and without the categorical dimensions of X.
+As removing those dimensions does not change significantly the error, we decided to compute the regression only with the 60 real-valued dimensions of $X$.
 
-Figure ~\ref{fig:reg_alphaRMSE} shows the train and test RMSE for a step size $\alpha$ varied from $10^{-3}$ to $10^{-0.5}$ with $20$ points in between.
+Figure ~\ref{fig:reg_alphaRMSE} shows train and test RMSE for a step size $\alpha$ varied from $10^{-3}$ to $10^{-0.5}$ with $20$ points in between.
 The convergence condition was reached when the variation of $\beta$ between two iterations was lower than $1.5 \alpha$.
 Cross-validation was made using $66 \% $ of the data as train data and the rest as test data.
-Train and test RMSE is optimal for $\alpha \approx 0.2$.
+Train and test RMSE are optimal for $\alpha \approx 0.2$.
 With higher $\alpha$ values, gradient descent is not able to converge and the error goes to $\infty$.
@@
 
-Figure ~\ref{fig:reg_lambdaRMSE} shows the train and test RMSE for $\lambda$ varied from $10^{-3}$ to $10^{7}$ with $60$ points in between.
+Figure ~\ref{fig:reg_lambdaRMSE} shows train and test RMSE for $\lambda$ varied from $10^{-3}$ to $10^{7}$ with $60$ points in between.
 Cross-validation was made using $66 \% $ of the data as train data and the rest as test data.
-We can see that $\lambda$ should be chosen between $0$ and $100$ to have the best prediction and that $\lambda$ higher than $1000$ increases the RMSE and the model under-fits.
+We can see that $\lambda$ should be chosen between $0$ and $100$ to have the best prediction and that $\lambda$ higher than $1000$ increases RMSE and the model under-fits.
 For high $\lambda$ values, the model and estimation bias are increased as their variance is reduced.
@@
 	\center
-	\includegraphics[width=2.6in]{figures/lambdaRMSE.png}
+	\includegraphics[width=2.8in]{figures/lambdaRMSE.png}
 	\caption{Ridge regression for a $66 \%$ - $33 \%$ split with 3 iterations for each $\lambda$.}
@@
 
-The median train/test RMSE are respectively equal to $526/539$, $551/562$ and $526/524$ for least-squares, least-squares GD and ridge regression.
-We observe that least-squares GD has a high bias and a low variance compared to least-squares that has a lower bias and a higher variance.
-Even if least-squares and ridge regression have similar train and test RMSE, we prefer to use ridge regression as its variance is lower and as its outliers are not very far from the median compared to those of least-squares.
+The medians of train/test RMSE are respectively equal to $526/539$, $551/562$ and $526/524$ for least-squares, least-squares GD and ridge regression.
+We observe that least-squares GD has a high bias and a low variance compared to least-squares.
+Even if least-squares and ridge regression have similar train and test RMSE, we prefer to use ridge regression as its variance is lower and as outliers are not very far from the median compared to outliers of least-squares.
 
@@
 Our classification-data contains $N=1500$ train-examples and $N=1500$ test-data.
-The train-data are composed of input and output variables $X$ and $y$.
+The train-data is composed of input and output variables $X$ and $y$.
 Each input $x_n$ has a dimensionality $D=37$ with 32 real-valued variables, 2 binary variables, 3 categorical variables with 5 categories.
@@
 Figures ~\ref{fig:class_dim23} and ~\ref{fig:class_dim35} show respectively the histogram of the $23^{th}$ and the $35^{th}$ variable of $X$.
-We see that the distribution is composed of a high peak and a Gaussian distribution.
-We observed that most of the real-valued dimensions have similar distributions and that data-points belonging to the peak or to the Gaussian in one dimensions are contained in the same part of the distribution in other dimensions.
+We see that the distribution is composed of a high peak and a sort of Gaussian distribution.
+We observed that most of the real-valued dimensions have similar distributions and that data-points belonging to the peak or to the Gaussian in one dimension are contained in the same part of the distribution in other dimensions.
 Moreover, the binary value of the $20^{th}$ dimension is different for data-points in the peak or in the Gaussian, so that we can use this dimension to separate the input data in two clusters.
 
-We also computed the eigenvalues of the matrix $X^{T}X$. As the smaller eigenvalue is equal to $58.4$, the matrix is full-rank and we do not need to use methods to lift the eigenvalues.
+We also computed the eigenvalues of the matrix $X^{T}X$. As the smaller eigenvalue is equal to $58.4$, the matrix is full-ranked and we do not need to use methods to lift the eigenvalues.
 
@@
 	\center
-	\subfigure[Histogram of the $23^{th}$ dimension of $X$.]{\includegraphics[width=2.1in]{figures/dim23.png} \label{fig:class_dim23}}
+	\subfigure[Histogram of the $23^{th}$ dimension of $X$.]{\includegraphics[width=2.2in]{figures/dim23.png} \label{fig:class_dim23}}
 	\hfill
-	\subfigure[Histogram of the $35^{th}$ dimension of $X$.]{\includegraphics[width=2.1in]{figures/dim35.png} \label{fig:class_dim35}}
+	\subfigure[Histogram of the $35^{th}$ dimension of $X$.]{\includegraphics[width=2.2in]{figures/dim35.png} \label{fig:class_dim35}}
 	\caption{Example of histogram to show the distribution of $X$ dimensions.
@@
 \subsection{Feature Transformations}
-The train-data are first separated in two clusters given the value of the 20th dimension of $X$.
-Each group of data are then treated separately.
+The train-data is first separated in two clusters given the value of the 20th dimension of $X$.
+Each group of data is then treated separately.
 The real-valued variables of $X$ are first normalised to have zero-mean and unitary standard deviation.
@@
 
-After the clustering of the test-data, each group is also treated separately. The real-valued input data are normalised given the means and standard deviations computed during the training.
+After the clustering of the test-data, each group is also treated separately. The real-valued input dimensions are normalised given the means and standard deviations computed during the training.
 
@@
 To improve the convergence, the step size was adapted at each time step to be proportional to the log likelihood such that $\alpha = log(\ell) * 10^{-4}$.
-The convergence criterion was reached when the difference between $\beta$ between two consecutive iterations was lower than $0.001$.
+The convergence criterion was reached when the difference between $\beta$ computed during two consecutive iterations was lower than $0.001$.
 
-Figure ~\ref{fig:class_penLogistic} shows the train and test loss function for $\lambda$ varied from $10^{-3}$ to $1$.
+Figure ~\ref{fig:class_penLogistic} shows train and test loss function for $\lambda$ varied from $10^{-3}$ to $1$.
+Cross-validation was made using $50\%$ of the data as train data and the rest as test data.
 We observe that changing $\lambda$ does not really improve the train and test loss function.
@@
 	\center
-	\includegraphics[width=2.6in]{figures/lambdaLossBox.png}
+	\includegraphics[width=2.8in]{figures/lambdaLossBox.png}
 	\caption{Penalised logistic regression for a $50 \% - 50\%$ split with 3 iterations for each $\lambda$.}
@@
 
+Figure ~\ref{fig:class_compare} compares two classification methods with $\lambda = 0.01$ for the penalised logistic regression.
+The medians of train/test loss function are respectively $0.061/0.126$ and $0.057/0.115$ for logistic regression and for the penalised method.
+We observe that penalised logistic regression gives smaller error and has lower variance than simple logistic regression.
+Even if the improvement is not really significant, we prefer to use the penalised method, as the other one must sometimes be stopped before a complete convergence because the cost function suddenly explodes to infinity.
+
+\begin{figure}[!h] % !t
+	\center
+	\includegraphics[width=3.0in]{figures/compareClassification.png}
+	\caption{Boxplots of train and test loss function for logistic regression (Log) and penalised logistic regression (PenLog) with 10 iterations for each method.}
+	\label{fig:class_compare}
+\end{figure}
+
 \section{Summary}
-In this report, we analyse a regression and a classification dataset.
+In this report, we analysed a regression and a classification dataset.
 We used k-means clustering and PCA to preprocess the regression data.
-We found that ridge regression give a reasonable fit with a test RMSE equal to $524 \pm 20$.
+We found that ridge regression gives a reasonable fit with a test RMSE equal to $524 \pm 20$.
 Least-squares and least-squares with gradient descent predictions have respectively a higher variance and a higher bias compared to ridge regression.
 
-We separated the classification data using a binary dimension that classify the data in two different distributions along the variable of the input data. We found that TODO.
+We separated the classification data using a binary dimension that classify the data in two different distributions along the variables of the input data.
+We found that penalised logistic regression gives a reasonable classification with a test loss function equal to $ 0.115 \pm 0.02$ and is more reliable than logistic regression.
 
